### 2007-05-27 talk/dyla2007/dyla.tex
Review by exarkun.  A couple of open issues, intermediate check-in.

\documentclass{llncs}

\usepackage{makeidx}
\usepackage{graphicx}

\begin{document}

\pagestyle{headings}

\title{How to \emph{not} write Virtual Machines for Dynamic Languages}

%\titlerunning{XXX}     % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used

\author{Carl Friedrich Bolz and Armin Rigo}

\authorrunning{Bolz and Rigo}   % abbreviated author list (for running head)

%%%% modified list of authors for the TOC (add the affiliations)
\tocauthor{Bolz and Rigo (D\"usseldorf)}

\institute{ Lehrstuhl Softwaretechnik und Programmiersprachen\\
Institut f\"{u}r Informatik, Universit\"at D\"usseldorf, Germany\\
  \email{ cfbolz@gmx.de, arigo@tunes.org}}

\maketitle

\begin{abstract}

We argue in this paper that one should not write interpreters for dynamic
languages manually but rather use meta-programming techniques and raise the
overall level at which they are implemented. We believe this to be
ultimately a better investment of efforts than the development of more and more advanced
general-purpose object oriented VMs.

\medskip

Dynamic languages are traditionally implemented by writing a virtual machine
centered around an interpreter and/or a built-in compiler and providing
the object model and memory management. When a language becomes more popular,
the limitations of such an implementation lead to the emergence of alternative
implementations that try to solve some of the problems. Another reason for new
implementations is the desire to have the language integrate well with existing,
well-tuned object-oriented virtual machine like the Java Virtual Machine. In this paper, we
describe the mechanisms that lead to an abundance of implementations and
explore some of the limitations of standard VMs.  We propose a
complementary alternative to writing VMs by hand and dealing with low-level
details, validated by the PyPy project: flexibly generating virtual machines
from a single abstract language ``specification'', inserting features and
low-level details automatically -- including good just-in-time compilers tuned
to the dynamic language at hand.
\footnote{This research was partially supported by the EU funded
 project: IST 004779 PyPy (PyPy: Implementing Python in Python).} \\

\end{abstract}

\section{Introduction}

Dynamic languages are traditionally implemented by writing a virtual
machine for them in a low-level language like C or in a language that
can relatively easily be turned into C.  The machine implements an
object model supporting the high level dynamic language's objects.  It
typically provides features like automatic garbage collection.  Recent
languages like Python, Ruby, Perl and JavaScript have complicated
semantics which are most easily mapped to a simple interpreter operating
on syntax trees or bytecode; simpler languages like Lisp and Self
typically have more efficient implementations based on code
generation.

The effort required to build a new virtual machine is relatively
large.  This is particularly true for languages which are complex
and in constant evolution. Language implementation communities from an
open-source or academic context have only limited resources. Therefore they
cannot afford to have a highly complex implementation and often choose simpler
techniques even if that entails lower execution speed. Similarly, fragmentation
(for example because of other implementations of the same language) is a
problem because it divides available resources. All these points also apply to
the implementation of domain-specific languages where it is important to keep
the implementation effort small.

For these reasons writing a virtual machine in C is problematic because it
forces the language implementer to deal with many low-level details (like
garbage collection and threading issues). Limitations
of the C implementation lead to alternative implementations which draw
resources from the reference implementation. An alternative to writing
implementations in C is to build them on top of one of the newer object oriented
virtual machines (``OO VM'') such as the JVM or the CLR. This is often wanted by
the community anyway, since it leads to the ability to re-use the libraries of
these platforms. However, if a C implementation existed before the
implementation of such a VM is started, this enters into conflict with the goal of
having to maintain essentially a single, simple enough implementation for a
given programming language: as the language becomes popular, there will be a
demand to have it run on various platforms -- high-level VMs as well as
C-level environments.

In this paper, we will argue that it is possible to
benefit from and integrate with OO VMs while keeping the dynamic
language implemented with a single, simple source code base.  The idea is
to write an interpreter for that language in another sufficiently
high-level but less dynamic language.  This interpreter plays the role
of a specification for the dynamic language.  With a sufficiently capable
translation toolchain we can then generate whole virtual machines from
this specification -- either wholely custom VMs for C-level operating
systems or VMs layered on top of various OO VMs.  In other words,
meta-programming techniques can be used to successfully replace a
foreseeable one-VM-fits-all standardization attempt.

The crux of the argument is that VMs for dynamic languages should not be
written by hand!  The PyPy project is the justification,
proving that the approach is
feasible in practice.  Just as importantly, it also brings new insights
and concrete benefits in term of flexibility and performance that go
beyond the state of the art.

In section \ref{sect:approaches} we will explore the way VMs are typically
implemented in C and on top of OO VMs and some of the problems of these
approaches, using various Python implementations as the main example. In
section \ref{sect:metaprogramming} we will describe our proposed
meta-programming approach.

\section{Approaches to Dynamic Language Implementation}
\label{sect:approaches}

\def\implname#1{\emph{#1}}

The observation that limitations of a C-based implementation of a dynamic
language leads to the emergence of additional implementations is clear
in the case of Python.  The reference implementation, \implname{CPython}
\cite{cpy251}, is a simple recursive interpreter.  \implname{Stackless
Python} \cite{stackless} is a fork that adds micro-threading
capabilities to Python. One of the reasons for not incorporating it back
into CPython was that it was felt that this would make the
implementation too complex. Another implementation of the Python
language is \implname{Psyco} \cite{psyco-software}, which adds a
JIT-compiler to CPython.  Finally, \implname{Jython} is a
re-implementation for the Java VM and \implname{IronPython} for
the CLR.  All of these need to be kept in sync with the relatively fast
evolution of the language.

With the emergence of the CLR and the JVM as interesting language
implementation platforms, it is sometimes argued that
communities should only develop an implementation of their language
for one of these platforms (preferably the argument author's favourite
one).

\subsection{Assessing the Advantages of Implementing a Language on Top of OO
VMs}

Implementing a language on top of an existing OO VM is in many ways easier than
implementing it in C. Let's take a look at the advantages that are usually
cited for basing a
language implementation of a dynamic language on a standard object oriented
virtual machine, for example the JVM or the CLR.

\begin{itemize}
\item
\emph{Better interoperability than the C level:} Since the VM offers a standard
object model and all the languages implemented on top of it are using it, it is
easier to integrate the languages that are running on top of the VM. This
allows reuse of libraries between all the implemented languages. This is
typically the most important reason to want an implementation on the VM in
the first place.

\item
\emph{Cross-platform portability:} Only the underlying VM has to be ported to
various hardware architectures and operating systems. The languages implemented
on top of it can then be run without change in various environments.

\item
\emph{Better tools:} Better IDEs, debuggers and profilers.

\item
\emph{Better implementation of low-level issues like garbage collection,
threading:} Since an OO VM is expected to be widely used and usually backed by
a company, it becomes worthwhile and possible to spend a lot of effort tuning
its garbage collector, threading model, exception support and other low-level
implementation details.

\item
\emph{Better performance:} Similarly, object-oriented VMs usually come with a
highly tuned just-in-time compiler to make them perform well without requiring
ahead-of-time compilation to machine language. This, in addition to the
previous point, leads to much better performance of the languages running on top
of the VM.

\item
\emph{Ease of implementation:} The implementation of a language on top of an OO
VM is easier because it starts at a higher level than C. Usually a
high-level language like Java or C\# is used for the language implementation,
both of which offer the language implementer a much higher level of abstraction
than when implementing in C.

\item
\emph{A single unified implementation base:} The CLR and JVM are trying
to position themselves as all-encompassing platforms; if one succeeds,
implementations of the dynamic language for other platforms might no longer
be required.
\end{itemize}

\noindent
The central theme of the benefits of OO VMs is the ability to
implement certain hard things only once and share the benefits between
all language implementations on top of the OO VM.  At a closer look,
some of these advantages are only partially true in practice:

\begin{itemize}
\item
\emph{Better performance:} So far it seems that performance of highly dynamic
languages is not actually significantly improved on OO VMs. 
Jython is around 5
times slower than CPython, for IronPython (which
gives up on at least one feature -- frame objects -- to improve performance)
the figures vary but it is mostly
within the same order of magnitude as CPython. The most important reason for
this is that the VM's JIT compilers are optimized for specific usage patterns
that are common in the primary language of the OO VM. To achieve good speeds, the
language implementers would have to carefully produce code that matches these
usage patterns, which is not a simple task.

\item
\emph{Better GCs:} While this is obvious in theory, OO VMs tend to have a much
higher memory overhead to start with (XXX ref)

\item
\emph{Cross-platform portability:} While this is true to some extent, the
situation with regard to portability is not significantly improved compared to
e.g.  C/POSIX, which is relatively portable as well. Also, portability sometimes
comes at the price of performance, because even if the OO VM is running on a
particular hardware architecture it is not clear that the JIT is tuned for this
architecture (or working at all), leading to significantly reduced
speed.

\item
\emph{Ease of implementation:} This point is disputable. On the one hand, OO
VMs typically allow the language implementer to start at a higher level. On the
other hand, they also enforce a specific object and execution model. This means
that the concepts of the implemented language need to be mapped to the
execution model of the underlying VM, which may or may not be easy, depending very
much on the language in question.

An example where this mapping does not work very well is Prolog. While there
exist several implementations of Prolog on top of the JVM \cite{prologcafe}
\cite{InterProlog} and one on .NET \cite{psharp},
they are not particular efficient, especially when compared to good Prolog VMs
written in C. This is mostly because the Prolog execution model, which
involves backtracking and deep recursion, does not fit the JVM and .NET very
well. Therefore the Prolog implementations on top of OO VMs resort to models
that are quite unnatural both for the OO VM and for Prolog.

Another important point that makes implementation of languages on top of OO VMs
harder is that typically OO VMs don't support meta-programming very well, or do so only
at the bytecode level.
\end{itemize}

Nevertheless, some of the benefits are real and very useful, the most
prominent of which being easy interaction with the rest of the VM. Furthermore, there
is better tool support and better GCs. Also, for languages where the execution
model fits the OO VM well, many of the disadvantages disappear.

\subsection{The Cost of Implementation-Proliferation}

The described proliferation of language implementations is a large problem for
language communities. Although most individual implementations exist for good
reasons, the sum of all of them and the need to keep them synchronized with the
reference implementations leads to a significant amount of duplicated work and division of
effort; this is especially true for open source languages which tend to evolve
quickly. At any one point in time some of the implementations will lag behind
which makes writing code which can work on all of the implementations harder.

Implementing a language on top of a OO VM has many advantages, so some
people propose the solution of standardizing on one particular OO VM to not have
to maintain implementations for several of them. While this would, in theory,
alleviate the problem it is unlikely to happen. On the one hand, many political
issues are involved in such a decision. On the other hand, deciding on a single
object and execution model would not be an equally good fit for all languages.

In the next section, we explore a different approach for implementing
dynamic languages that we hope is able to solve many of the problems of
implementing a language, in particular the problem of an explosion of the number
of implementations.

\section{Meta-Programming Is Good}
\label{sect:metaprogramming}

The present paper proposes to approach the implementation of dynamic
languages from a meta-level: virtual machines for such languages should
not be written by hand, but generated automatically ``around'' an
interpreter playing the role of a high-level description of the language.  We
argue that this approach gives many of the benefits usually expected by
an implementer when he decides to target an existing object-oriented
virtual machine.  It also gives other benefits that we will describe --
mostly in term of flexibility.  Most importantly, it lets a
community write a single source implementation of the language, avoiding
the time-consuming task of keeping several of them in sync.  The single
source can be used to generate either custom VMs for C-like
environments or interpreters running on top of OO VMs.  This makes it
practical to experiment with large changes to the language and with
entirely new languages, such as domain-specific languages, while at any
time being able to run the implemented language in a variety of
environments, from C/Posix to the JVM to .NET.

\subsection{PyPy architecture}

We implemented this idea in the PyPy project \cite{pypy}.  The dynamic language
for which we wrote an interpreter is Python.  It is a language which,
because of its size and rather intricate semantics, is a good target for
our approach in the following sense: its previous reimplementations
(Jython for the JVM and IronPython for .NET) have each proved to be very
time-consuming to maintain.  Our implementation is, by construction,
easier to maintain and extremely portable (including to C/Posix, to the
JVM and to .NET).

In meta-programming terms, the PyPy architecture is as follows:

\begin{itemize}

\item
We use a very expressive \emph{object language} (RPython -- an analyzable
subset of Python) as the language in which the complete Python
interpreter is written, together with the implementation of its
built-in types.  The language is still close to Python, e.g.  it is
object-oriented, provides rich built-in types and has automatic memory
management.  In other words, the source code of our complete Python
interpreter is mostly free of low-level details -- no explicit memory
management, no pieces of C or C-level code.

\item
We use a very expressive metalanguage (namely regular Python) to
perform the analysis of RPython code (control flow and data flow
construction, type inference, etc.) and its successive
transformations.

\item
This meta-programming component of PyPy is called the \emph{translation
framework}, as it translates RPython source code (i.e. the full Python
interpreter) into lower-level code.  Its purpose is to add aspects to
and specialize the interpreter to fit a selectable virtual or hardware
runtime environment.  This either turns the interpreter into a
standalone virtual machine or integrates it into an existing OO VM.
The necessary support code -- e.g. the garbage collector when
targeting C -- is itself written in RPython in much the same spirit
that the Jikes RVM's GCs are written in Java \cite{JikesGC}; as needed, it is
translated together with the interpreter to form the final custom VM.
\end{itemize}

A detailed description of this translation process is beyond the scope of the
present paper; it can be found in \cite{pypyvmconstruction}.  The actual Python
interpreter of PyPy and the results we achieved by translating it to C, LLVM
\cite{LLVM} and .NET are described in \cite{architecture} \cite{translationdoc}
These results show that the
approach is practical and gives results whose performance is within the same
order of magnitude (within a factor of 2 and improving) of the hand-written,
well-tuned CPython, the C reference implementation.  These figures do not
include the spectacular speed-ups obtained in some cases by the JIT compiler
described in section \ref{subsect:dynamic_compilers}.

In the sequel (??? - what is the sequel), we will focus on the relative advantages and
inconveniences of the PyPy approach compared to the approach of
hand-writing a language implementation on top of an OO VM.

\subsection{A single source}

Our approach -- a single ``meta-written'' implementation -- naturally
leads to language implementations that have various advantages over the
``hand-written'' implementations.  Firstly, it is a single-source
approach -- we explicitly seek to solve the problem of proliferation of
implementations.  In the sequel (???), we will show that this goal can be
achieved without giving up the advantages of hand-written
implementations for OO VMs.  Moreover, there are additional advantages which,
in our opinion, are significant enough to hint that meta-programming,
though not widely used in general-purpose programming, is an essential
tool in a language implementer's toolbox.

\subsection{Writing the interpreter is easier}

A first point is that it makes interpreters easy to write, update and
generally experiment with.  More expressiveness helps at all levels: our
Python interpreter is written in RPython as a relatively simple
interpreter and is, in some respects, easier to understand than CPython.  We are
using its high level and flexibility to quickly experiment with features
or implementation techniques in ways that would, in a traditional
approach, require pervasive changes to the source code.  For example,
PyPy's Python interpreter can optionally provide lazily computed objects
-- a 150-lines extension in PyPy that would require global changes in
CPython.  Further examples can be found in our technical reports; we
should notably mention an extension adding a state-of-the-art security
model for Python based on data flow tacking \cite{D12.1}, and general
performance improvements found by extensive experimentation \cite{D06.1}, some
of which were back-ported to CPython.

\subsection{Separation of concerns}

At the level of the translation framework, the ability to change or
insert new whole-program transformations makes some aspects of the
interpreter easier to deal with.  By ``aspect'' we mean, in the original
AOP sense, a feature that is added to an object program by a
meta-program.  The most obvious example in our context is the insertion
of a garbage collector (chosen among several available ones) for the
target environments that lack it.  Another example is the translation of
the interpreter into a form of continuation-passing style (CPS), which
allows the translated interpreter to provide coroutines even though its
source is written in a simple highly recursive style.  For more details
and other examples of translation-level transformations, see
\cite{D07.1}.

%A more subtle example of separation of concerns is the way our generated
%implementations can be integrated with a host OO VM.  As mentioned
%above, an implementer deciding to directly target a specific OO VM needs
%not only good knowledge of the OO VM in question and its object model --
%he must fit the language into the imposed models.  Instead, in our
%approach this task is done at two levels: in a first step, a stand-alone
%interpreter is written -- which, if translated to a given OO VM, would
%simply give an interpreter for the dynamic language which is unable to
%communicate with the host VM.  Integration comes as a second step, and
%occurs at a different level, by introducing mappings between the
%relevant classes of the interpreter and the corresponding classes of the
%OO VM.

\subsection{The effort of writing a translation toolchain}

What are the efforts required to develop a translation toolchain capable
of analyzing and transforming the high-level source code and generating
lower-level output in various languages?

Although it is able to generate, among other things, a complete, custom
VM for C-like environments, we found that the required effort that must
be put into the translation toolchain was still much lower than that of
writing a good-quality OO VM.  A reason is that a translation toolchain
operates in a more static way, which allows it to leverage good C
compilers.  It is self-supporting: pieces of the implementation can be
written in RPython as well and translated along with the rest of the
RPython source, and they can all be compiled and optimized by the C
compiler.  In order to write an OO VM in this style you need to start by
assuming an efficient dynamic compiler.

Of course, the translation toolchain, once written, can also be reused
to implement other languages, and possibly tailored on a case-by-case
basis to fit the specific needs of a language.  The process is
incremental: we can add more features as needed instead of starting from
a maximal up-front design, and gradually improve the quality of the
tools, the garbage collectors, the various optimizations, etc.

Writing a good garbage collector remains hard, though.  At least, it is
easy to experiment with various kind of GCs, so we started by just using
the conservative Boehm \cite{Boehm} collector for C and moved up to a
range of simple custom collectors -- reference counting, mark-and-sweep,
etc.  Ultimately, though, more advanced GCs will be needed to get the
best performance.  It seems that RPython, enhanced with support for
direct address manipulations, is a good language for writing GCs, so it
would be possible for a GC expert to write one for our translation
framework.  However, this is not the only way to obtain good GCs:
existing GCs can also be reused.  Good candidates are the GCs written in
the Jikes RVM \cite{JikesGC}.  As they are in Java, it should be
relatively straightforward to add a translation step that turns one of
them into RPython (or directly into our RPython-level intermediate
representation) and integrates it with the rest of the program being
translated.

In summary, developing a meta-programming translation toolchain requires
work, but it can be done incrementally, it can reuse existing code, and
it results in a toolchain that is itself highly reusable and flexible in
nature.

\subsection{Dynamic compilers}
\label{subsect:dynamic_compilers}

As mentioned above, the performance of the VMs generated by our
translation framework are quite acceptable -- e.g. the Python VM
generated via C code is much faster than Jython running on the best
JVMs.  Of course, the JIT compilers in these JVMs are essential to
achieve even this performance, which further proves the point that
writing good OO VMs -- especially ones meant to support dynamic
languages -- is a lot of work.

The deeper problem with the otherwise highly-tuned JIT compilers of the
OO VMs is that they are not a very good match for running dynamic
languages.  It might be possible to tune a general-purpose JIT compiler
enough and write the dynamic language implementation accordingly so
that most of the bookkeeping work involved in running the dynamic
language can be removed -- dispatching, boxing, unboxing...  However
this has not been demonstrated yet.

By far the fastest Python implementation, Psyco \cite{psyco-software}, contains a
hand-written language-specific dynamic compiler.  PyPy's translation
tool-chain is able to extend the generated VMs with an automatically
generated dynamic compiler that uses techniques similar to those of Psyco
\cite{Psyco-paper}, derived from the
interpreter.  This is achieved by a pragmatic application of partial
evaluation techniques guided by a few hints added to the source of the
interpreter.  In other words, it is possible to produce a reasonably
good language-specific JIT compiler and insert it into a VM, alongside
with the necessary support code and the rest of the regular interpreter.

This result was one of the major goals and motivations for the whole
approach.  By construction, the JIT does not get out of sync when the
language evolves, and any code written in the dynamic language runs
correctly under the JIT.  Some very simple Python examples run more than
100 times faster.  At the time of this writing this is still rather
experimental, and the techniques involved are well beyond the scope of
the present paper.  The reader is referred to \cite{D08.2} for more
information.

\section{Conclusion}

Here are the two central points that we have asserted in the present
paper:

%XXX doesn't look entirely nice
\begin{itemize}
\item \emph{Do not write dynamic language implementations ``by hand''.}
Writing them more abstractly, at a higher level, has primarily only
advantages, among them the avoidance of a proliferation of implementations
growing out of sync.  Writing interpreters both flexibly and efficiently
is difficult and meta-programming is a good way to achieve it.
Moreover, this is not incompatible with targetting and benefiting from
existing high-quality object-oriented virtual machines like those of the Java and .NET.

\item \emph{Do not write VMs ``by hand''.}
Writing language-specific virtual machines is a time-consuming task for
medium to large languages.  Unless large amounts of resources can be
invested, the resulting VMs are bound to have limitations which lead to
the emergence of many implementations, a fact that is taxing precisely
for a community with limited resources.  (This is of course even more
true for general-purpose VMs.)
\end{itemize}

\noindent
As a better alternative, we advocate a more general usage of
meta-programming:

\begin{itemize}
\item \emph{Let's write more meta-programming translation toolchains.}
Aside from the advantages described in section
\ref{sect:metaprogramming}, a translation toolchain need not be
standardized for inter-operability but can be tailored to the needs of
each project.  Diversity is good; there is no need to attempt to
standardize on a single OO VM.
\end{itemize}

The approach we outlined is actually just one in a very large, mostly
unexplored design space; it is likely that some of the choices made in
PyPy will turn out to be suboptimal.  We are hoping that other
toolchains will emerge over time, exploring other aspects and proposing
other solutions.  By their ``meta'' nature, these multiple approachs
should be easier to bridge together than, say, multiple OO VMs with
different object and runtime models.  We believe that further research
in this area might open the door to better solutions for
interoperability in general -- e.g. high-level bridges instead of
(virtual-)machine-level ones, enabled by cross-translation.

We believe this to be ultimately a better investment of efforts than the
development of more advanced general-purpose OO VMs.

% ---- Bibliography ----
%\begin{small}
\bibliographystyle{abbrv}
\bibliography{dyla}
%\end{small}

\end{document}