### 2008-12-02 talk/ecoop2009/abstract.tex
first draft of the ecoop paper

\begin{abstract}
Dynamic languages are increasingly popular etc. etc.
\end{abstract}