### 2008-12-19 talk/ecoop2009/conclusion.tex
add (empty) conclusion file

\section{Related Word}
XXX add me

\section{Conclusion}

XXX add me