### 2008-12-19 talk/ecoop2009/conclusion.tex
add material to the related work section, needs an overhaul

\section{Related Word}
Promotion is a concept that we have already explored in other contexts. Psyco is
a run-time specialiser for Python that uses promotion (called ``unlift'' in
\cite{DBLP:conf/pepm/Rigo04}). However, Psyco is a manually written JIT, is
not applicable to other languages and cannot be retargetted.

 - reference standard jit stuff, PICs \cite{hoelzle_optimizing_1991}, this one: http://www.cs.ucsb.edu/~urs/oocsb/self/papers/type-feedback.html

 - compare against tracing JITs \cite{gal_hotpathvm_2006}. tracing jits make
 implementing jits much easier, maybe less need for JIT generation. also, they
 concentrate on loops, which makes them produce a lot less code (which will need
 to be addressed in our work as well).

 so far tracing JITs have less general allocation removal techniques, which
 makes them get less speedup in a dynamic language with boxing

in my master I had this snippet:

"""
The generated code typically contains guards; in recent research
\cite{gal_incremental_2006} on Java, these guards' behaviour is extended to be
similar to our promotion.  This has been used twice to implement a dynamic
language (JavaScript), by Tamarin\footnote{{\tt
http://www.mozilla.org/projects/tamarin/}} and in \cite{chang_efficient_2007}.
"""

There has been an enormous amount of work on partial evaluation for compiler
generation. A good introduction is given in \cite{Jones:peval}. However, most of
it is for generating ahead-of-time compilers, which cannot produce very good
performance results for dynamic languages.

However, there is also some research on runtime partial evaluation. One of the
earliest examples is Tempo for C
\cite{DBLP:conf/popl/ConselN96,DBLP:conf/dagstuhl/ConselHNNV96}. However, it is
essentially an offline specializer ``packaged as a library''; decisions about
what can be specialized and how are pre-determined.

Another work in this direction is DyC \cite{grant_dyc_2000}, another runtime
specializer for C.  Specialization decisions are also pre-determined, but
``polyvariant program-point specialization'' gives a coarse-grained equivalent
of our promotion.  Targeting the C language makes higher-level specialization
difficult, though (e.g.\ \texttt{mallocs} are not removed).

Greg Sullivan introduced "Dynamic Partial Evaluation", which is a special
form of partial evaluation at runtime \cite{sullivan_dynamic_2001} and describes
an implementation for a small dynamic language based on lambda calculus. This
work is conceptually very close to our own.
% XXX there are no performance figures, we have no clue how much of this is
% implemented. not sure how to write this

 - apart from that we maybe should talk about escape analysis
 \cite{Blanchet99escapeanalysis}, \cite{Choi99escapeanalysis}
 our algorithm is totally simple-minded in comparison, but is useful in practise

\section{Conclusion}

XXX add me