### 2008-12-21 talk/ecoop2009/conclusion.tex
add conclusion and future work

\section{Related Work}

Promotion is a concept that we have already explored in other contexts. Psyco is
a run-time specialiser for Python that uses promotion (called ``unlift'' in
\cite{DBLP:conf/pepm/Rigo04}). However, Psyco is a manually written JIT, is
not applicable to other languages and cannot be retargetted.

Moreover, the idea of promotion of is a generalization of \emph{Polymorphic
  Inline Caches} \cite{hoelzle_optimizing_1991}, as well as the idea of using
runtime feedback to produce more efficient code
\cite{hoelzle_type_feedback_1994}.

PyPy-style JIT compilers are hard to write manually, thus we chose to write a
JIT generator.  Tracing JIT compilers \cite{gal_hotpathvm_2006} also gives
good results but are much easier to write, making the need for an automatic
generator less urgent.  However so far tracing JITs have less general
allocation removal techniques, which makes them get less speedup in a dynamic
language with boxing.  Another difference is that tracing JITs concentrate on
loops, which makes them produce a lot less code.  This issue will be addressed
by future research in PyPy.

The code generated by tracing JITs code typically contains guards; in recent research
\cite{gal_incremental_2006} on Java, these guards' behaviour is extended to be
similar to our promotion.  This has been used twice to implement a dynamic
language (JavaScript), by Tamarin\footnote{{\tt
http://www.mozilla.org/projects/tamarin/}} and in \cite{chang_efficient_2007}.

There has been an enormous amount of work on partial evaluation for compiler
generation. A good introduction is given in \cite{Jones:peval}. However, most of
it is for generating ahead-of-time compilers, which cannot produce very good
performance results for dynamic languages.

However, there is also some research on runtime partial evaluation. One of the
earliest examples is Tempo for C
\cite{DBLP:conf/popl/ConselN96,DBLP:conf/dagstuhl/ConselHNNV96}. However, it is
essentially an offline specializer ``packaged as a library''; decisions about
what can be specialized and how are pre-determined.

Another work in this direction is DyC \cite{grant_dyc_2000}, another runtime
specializer for C.  Specialization decisions are also pre-determined, but
``polyvariant program-point specialization'' gives a coarse-grained equivalent
of our promotion.  Targeting the C language makes higher-level specialization
difficult, though (e.g.\ \texttt{mallocs} are not removed).

Greg Sullivan introduced "Dynamic Partial Evaluation", which is a special
form of partial evaluation at runtime \cite{sullivan_dynamic_2001} and describes
an implementation for a small dynamic language based on lambda calculus. This
work is conceptually very close to our own.
% XXX there are no performance figures, we have no clue how much of this is
% implemented. not sure how to write this

Our algorithm to avoid allocation of unneeded intermediate objects fits into
the research area of escape analysis: in comparison to advanced techniques
\cite{Blanchet99escapeanalysis}, \cite{Choi99escapeanalysis} our algorithm is
totally simple-minded, but it is still useful in practise.

\section{Conclusion and Future Work}

In this paper we presented PyPy's JIT compiler generator, based on partial
evaluation techniques, which can automatically turn an interpreter into a JIT
compiler, requiring the language developers to only add few \texttt{hint}s to
guide the generation process.

We showed that classical partial evaluation cannot remove all the overhead
proper of dynamically typed languages, and how the new operation called
\emph{promotion} solves the problem, by delaying compile-time until the JIT
knows enough to produce efficient code, and by continuously intermixing
compile-time and runtime.  Moreover, we showed that our simple but still
practically useful technique to avoid allocation of intermediate unnecessary
objects plays well with promotion and helps to produce even better code.

Finally, we presented the CLI backend for PyPy's JIT compiler generator, whose
goal is to produce .NET bytecode at runtime.  We showed how it is possible to
circumvent intrinsic limitations of the virtual machine to implement
promotion.  As a result, we proved that the idea of \emph{JIT layering} is
worth of further exploration, as it makes possible for dynamically typed
languages to be even faster than their statically typed counterpart in some
circumstances.

As a future work, we want to explore different strategies to make the frontend
producing less code, maintaining comparable or better performances.  In
particular, we are working on a way to automatically detect loops in the user
code, as tracing JITs do \cite{gal_hotpathvm_2006}.  By compilining whole
loops at once, the backends should be able to produce better code than today.

At the moment, some bugs and minor missing features prevent the CLI JIT
backend to handle more complex languages such as Python and Smalltalk.  We are
confident that once these problems will be fixed, we will get performance
results comparable to TLC, as the other backends already demonstrate
\cite{PyPyJIT}.  Moreover, if the current implementation of flexswitches will
prove to be too slow for some purposes, we want to explore alternative
implementation strategies, also considering the new features that might be
integrated into virtual machines.