### 2009-01-19 talk/ecoop2009/tutorial/proposal.tex
(cfbolz, antocuni, arigo): expanded the proposal

\documentclass{llncs}

 %%%% Compression Light+: LNCS margin reduced by +/-7mm along all edges (RG).
%\textwidth=130mm   % LNCS: 122mm
%\textheight=203mm  % LNCS: 193mm

\renewcommand{\baselinestretch}{0.97}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[sans]{dsfont}
\usepackage{color}
\usepackage{ifthen}
\usepackage{xspace}
\usepackage{listings}
\usepackage{fancyvrb}
\usepackage{multirow}
\usepackage[pdftex]{graphicx}

%\input{macros}

\pagestyle{plain}

%\lstset{mathescape=true,language=Java,basicstyle=\tt,keywordstyle=\bf}
\lstset{language=Python,
        basicstyle=\scriptsize\ttfamily,
        keywordstyle=\color{blue}, % I couldn't find a way to make chars both bold and tt
        frame=none,
        stringstyle=\color{blue},
        fancyvrb=true,
        xleftmargin=20pt,xrightmargin=20pt,
        showstringspaces=false}

\setlength{\tabcolsep}{1ex}

%\renewcommand{\baselinestretch}{.98}
\newboolean{showcomments}
\setboolean{showcomments}{false}
\ifthenelse{\boolean{showcomments}}
  {\newcommand{\nb}[2]{
    \fbox{\bfseries\sffamily\scriptsize#1}
    {\sf\small$\blacktriangleright$\textit{#2}$\blacktriangleleft$}
   }
   \newcommand{\version}{\emph{\scriptsize$-$Id: main.tex 19055 2008-06-05 11:20:31Z cfbolz $-$}}
  }
  {\newcommand{\nb}[2]{}
   \newcommand{\version}{}
  }

\newcommand\dacom[1]{\nb{DA}{#1}}
\newcommand\cfbolz[1]{\nb{CFB}{#1}}
\newcommand\anto[1]{\nb{ANTO}{#1}}
\newcommand\arigo[1]{\nb{AR}{#1}}
\newcommand{\commentout}[1]{}

\let\oldcite=\cite

\renewcommand\cite[1]{\ifthenelse{\equal{#1}{XXX}}{[citation~needed]}{\oldcite{#1}}}

\begin{document}
\title{Tutorial Proposal: Writing Interpreters for Dynamic Languages Using PyPy
(and Getting Them Fast)}

\author{Carl Friedrich Bolz\inst{2} \email{cfbolz@gmx.de} \and Antonio Cuni\inst{1} \email{anto.cuni@gmail.com} \and Armin Rigo\inst{2} \email{arigo@tunes.org}}

\institute{DISI, University of Genova, Italy 
\and 
Softwaretechnik und Programmiersprachen
 Heinrich-Heine-Universit\"at D\"usseldorf}

\maketitle

The easiest way to implement a dynamic language such as JavaScript or Python is
to write an interpreter for it; however, interpreters are slow. An alternative
is to write a compiler; writing a compiler that targets a high level virtual
machine like CLI or JVM is easier than targeting a real CPU, but it still
require a lot of work, as IronPython, Jython, JRuby demonstrate. Moreover,
the various platforms have to be targeted independently.

The PyPy project\footnote{\url{http://codespeak.net/pypy}} \cite{RiBo07_223}
aims to make the implementation of dynamic
languages easier by providing a toolchain that allows to translate an
interpreter to various target platforms, including the JVM, .NET and C/Posix. 
In addition, we are working on being able to automatically transform the
interpreter into a specializing JIT-compiler to vastly increase performance.
The goal is to minimize the effort required to get a fast implementation of a
dynamic language.

This tutorial will give an insight into the usage of the PyPy translation
toolchain by giving a step-by-step introduction. The topics covered are how to
write an interpreter for a dynamic language for PyPy and make it possible to
automatically generate a JIT for it.  We will showcase a typical interpreter for a
small dynamic language as a running example.

Outline of the tutorial:
\begin{itemize}
 \item \textbf{Introduction to PyPy:} Introduce the PyPy project and give an
 overview of its goals.
 \item \textbf{Overview of the example language:} Present the small language and
 its interpreter that will be used as a running example throughout the tutorial.
 \item \textbf{Translation:} Show how to compile the interpreter to various
 platforms, including .NET and C/Posix.
 \item \textbf{JIT-Generation:} Show how to apply the JIT-generator to the
 interpreter to automatically get a JIT for the language.
\end{itemize}

There is a current resurgence in research on JIT-compilers for dynamic
languages, mostly focusing on JavaScript implementations. PyPy generates such
JITs so that they don't have to be hand-written. This requires much less effort
for the language implementor.

The presenters are part of the PyPy core development team and are all involved
in the ongoing research on the JIT generator.

\bibliographystyle{plain}
\bibliography{proposal}

\end{document}