### 2009-04-03 talk/icooolps2009-dotnet/paper.tex
Typos.

\documentclass{acm_proc_article-sp}

\usepackage{ifthen}
\usepackage{fancyvrb}
\usepackage{color}
\usepackage{ulem}
\usepackage{listings}

\lstset{language=Python,
        basicstyle=\scriptsize\ttfamily,
        keywordstyle=\color{blue}, % I couldn't find a way to make chars both bold and tt
        frame=none,
        stringstyle=\color{blue},
        fancyvrb=true,
        xleftmargin=10pt,xrightmargin=10pt,
        showstringspaces=false}

\newboolean{showcomments}
\setboolean{showcomments}{true}
\ifthenelse{\boolean{showcomments}}
  {\newcommand{\nb}[2]{
    \fbox{\bfseries\sffamily\scriptsize#1}
    {\sf\small$\blacktriangleright$\textit{#2}$\blacktriangleleft$}
   }
   \newcommand{\version}{\emph{\scriptsize$-$Id: main.tex 19055 2008-06-05 11:20:31Z cfbolz $-$}}
  }
  {\newcommand{\nb}[2]{}
   \newcommand{\version}{}
  }

\newcommand\davide[1]{\nb{DAV}{#1}}
\newcommand\cfbolz[1]{\nb{CFB}{#1}}
\newcommand\anto[1]{\nb{ANTO}{#1}}
\newcommand\arigo[1]{\nb{AR}{#1}}
\newcommand{\commentout}[1]{}

\normalem

\let\oldcite=\cite

\renewcommand\cite[1]{\ifthenelse{\equal{#1}{XXX}}{[citation~needed]}{\oldcite{#1}}}

\begin{document}

%% \title{Automatic generation of JIT compilers for dynamic languages in
%%   .NET\thanks{This work has been partially supported by MIUR EOS DUE -
%%     Extensible Object Systems for Dynamic and Unpredictable Environments and
%%     by the EU-funded project: IST 004779 PyPy (PyPy: Implementing Python in
%%     Python).}}

\title{Faster than C\#: efficient implementation of dynamic languages on
  .NET\thanks{This work has been partially supported by MIUR EOS DUE -
    Extensible Object Systems for Dynamic and Unpredictable Environments and
    by the EU-funded project: IST 004779 PyPy (PyPy: Implementing Python in
    Python).}}

%% Alternative title: Faster than C\#: the future of dynamic languages on .NET

\numberofauthors{3}
\author{
\alignauthor Antonio Cuni\\
       \affaddr{DISI, University of Genova}\\
       \affaddr{Italy}\\
       \email{cuni@disi.unige.it}
\alignauthor Davide Ancona\\
       \affaddr{DISI, University of Genova}\\
       \affaddr{Italy}\\
       \email{davide@disi.unige.it}
\alignauthor Armin Rigo\\
       \email{arigo@tunes.org}
}
\maketitle

\begin{abstract}
The CLI, i.e. the virtual machine at the core of the .NET environment, is a
statically typed virtual machine.  Although it is being used as a target for
many different languages, it is hard to provide highly efficient implementation
of languages whose model is too different than the one of the VM.  In
particular, dynamically typed languages are typically order of magnitude
slower than, e.g., C\#.

\anto{XXX: should we cite the DLR? How to prove that DLR programs are much
  slower than C\#?}

Recent developments show that JIT compilers can exploit runtime type
information to generate quite efficient code.  Unfortunately, writing a JIT
compiler is far from being simple.  In this paper we report our positive
experience with automatic generation of JIT compilers as supported by the PyPy
infrastructure, by focusing on JIT compilation for .NET.

JIT compilation for .NET means to generate .NET bytecode at runtime, which
will be further compiled down to machine code by .NET's own JIT compiler.  The
main and novel contribution of this paper is to show that this
\emph{two-layers JIT} technique is effective, and can give very high speedups,
up to making dynamic languages faster than C\# itself under certain
circumstances.

The practicality of the approach is demonstrated by showing some promising
experiments done with benchmarks written in a simple dynamic language.
\end{abstract}

\input{intro}
\input{clibackend}
\input{benchmarks}
\input{conclusion}

\bigskip

\bibliographystyle{abbrv}
\bibliography{paper}

\end{document}

% LocalWords:  JIT PyPy