### 2009-04-06 talk/icooolps2009-dotnet/conclusion.tex
adapt che conclusions to the new paper

\section{Related Work}

Flexswitches are closely related to the concept of \emph{promotion}, as
described by \cite{PyPyJIT}, \cite{PyPyJIT09}.
Psyco is
a run-time specialiser for Python that uses promotion (called ``unlift'' in
\cite{DBLP:conf/pepm/Rigo04}). However, Psyco is a manually written JIT, is
not applicable to other languages and cannot be retargetted.

Moreover, the idea of promotion is a generalization of \emph{Polymorphic
  Inline Caches} \cite{hoelzle_optimizing_1991}, as well as the idea of using
runtime feedback to produce more efficient code
\cite{hoelzle_type_feedback_1994}.

PyPy-style JIT compilers are hard to write manually, thus we chose to write a
JIT generator.  Tracing JIT compilers \cite{gal_hotpathvm_2006} also give
good results but are much easier to write, making the need for an automatic
generator less urgent.  However so far tracing JITs have less general
allocation removal techniques, which makes them get less speedup in a dynamic
language with boxing.  Another difference is that tracing JITs concentrate on
loops, which makes them produce a lot less code.  This issue is being addressed
by current research in PyPy \cite{PyPyTracing}.

The code generated by tracing JITs code typically contains guards; in recent research
\cite{gal_incremental_2006} on Java, these guards' behaviour is extended to be
similar to our promotion.  This has been used twice to implement a dynamic
language (JavaScript), by Tamarin\footnote{{\tt
http://www.mozilla.org/projects/tamarin/}} and in \cite{chang_efficient_2007}.

IronPython and Jython are two popular implementations of Python for,
respectively, the CLI and the JVM, whose approach differs fundamentally from
PyPy.  The source code of PyPy contains a Python interpreter, which the JIT
compiler is automatically generated from: the resulting executable contains
both the interpreter and the compiler, so that it is possible to compile only
the desired parts of the program.  On the other hand, both IronPython and
Jython implements only the compiler: both compile code lazily (when a Python
module is loaded), but since they do not exploit the extra information
potentially available at runtime, it is more a delayed static compilation than
a true JIT one.  As a result, they run Python programs much slower than their
equivalent written in
C\#\footnote{http://shootout.alioth.debian.org/gp4/\\benchmark.php?test=all\&lang=iron\&lang2=csharp}
or
Java\footnote{http://blog.dhananjaynene.com/2008/07/performance-comparison-c-java-python-ruby-jython-jruby-groovy/}.

The \emph{Dynamic Language Runtime}\footnote{http://www.codeplex.com/dlr}
(DLR) is a library designed to ease the implementation of dynamic languages
for .NET: IronPython is based on the DLR, so the same remarks apply.

\anto{XXX: please review the last two paragraphs}

\section{Conclusion and Future Work}

In this paper we gave an overview of PyPy's JIT compiler generator,
which can automatically turn an interpreter into a JIT
compiler, requiring the language developers to only add few \texttt{hint}s to
guide the generation process.

Then, we presented the CLI backend for PyPy's JIT compiler generator, whose
goal is to produce .NET bytecode at runtime.  We showed how it is possible to
circumvent intrinsic limitations of the virtual machine to implement
flexswitches.  As a result, we proved that the idea of \emph{JIT layering} is
worth of further exploration, as it makes possible for dynamically typed
languages to be even faster than their statically typed counterpart in some
circumstances.

As a future work, we want to explore different strategies to make the frontend
producing less code, maintaining comparable or better performances.  In
particular, we are working on a way to automatically detect loops in the user
code, as tracing JITs do \cite{gal_hotpathvm_2006}.  By compilining whole
loops at once, the backends should be able to produce better code than today.

At the moment, some bugs and minor missing features prevent the CLI JIT
backend to handle more complex languages such as Python and Smalltalk.  We are
confident that once these problems will be fixed, we will get performance
results comparable to TLC, as the other backends already demonstrate
\cite{PyPyJIT}.  Moreover, if the current implementation of flexswitches will
prove to be too slow for some purposes, we want to explore alternative
implementation strategies, also considering the new features that might be
integrated into virtual machines.

In particular, the \emph{Da Vinci Machine
  Project} \footnote{http://openjdk.java.net/projects/mlvm/} is exploring and
implementing new features to ease the implementation of dynamic languages on
top of the JVM: some of these features, such as the new
\emph{invokedynamic}\footnote{XXX} instruction and the \emph{tail call
  optimization} can probably be exploited by a potential JVM backend to
generate even more efficient code.