### 2010-02-07 talk/pycon2010/keynote/talk.tex
Stuff I did so far

\documentclass[utf8x, 14pt]{beamer}

\mode<presentation>
{
  \usetheme{Warsaw}
}

\usepackage[english]{babel}

\usepackage[utf8x]{inputenc}
\usepackage{tikz}

\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{color}

\setbeamerfont{big}{size*={64pt}{0pt}}

\title{The status of PyPy}

\author{Maciej Fijałkowski}

\institute[merlinux GmbH]
{ merlinux GmbH }

\date{Pycon 2010, February 20th 2010, Atlanta}

\begin{document}

\begin{frame}
  \titlepage
  \begin{figure}
    \begin{tabular}{c c c}
    \includegraphics[width=.30\textwidth]{pypy-logo.png}
    &
    \hspace{2cm}
    &
    \includegraphics[width=.25\textwidth]{merlinux-logo.png}
    \end{tabular}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Introduction}
  \begin{itemize}
    \item What we did in the last year?
    \item What we achieved?
    \item Plans for near future
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Last year highlits}
  \pause
  \usebeamerfont{big}
  JIT
\end{frame}

\begin{frame}
  \frametitle{February 2009}
  \begin{tikzpicture}[remember picture, overlay]
    \node [shift={(1cm, -4cm)}] at (current page.north west) {
      \begin{tikzpicture}[remember picture, overlay, font={\fontsize{8pt}{0pt} \ttfamily}]
        \node [shift={(-1cm, 1.4cm)}, above right] {
          \includegraphics[width=.15\textwidth]{pypy-logo.png}};
        \draw [thick, ->] (1.2cm, 2.5cm) -- (10cm, 2.5cm);
      \end{tikzpicture}
    };
  \end{tikzpicture}
  \begin{itemize}
    \item We merge experiments into mailine PyPy
    \item xxx
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{March 2009}
  \begin{tikzpicture}[remember picture, overlay]
    \node [shift={(1cm, -4cm)}] at (current page.north west) {
      \begin{tikzpicture}[remember picture, overlay, font={\fontsize{8pt}{0pt} \ttfamily}]
        \node [shift={(-1cm, 1.4cm)}, above right] {
          \includegraphics[width=.15\textwidth]{pypy-logo.png}};
        \draw [thick, ->] (1.2cm, 2.5cm) -- (10cm, 2.5cm);
        \draw [fill=blue!20,very thin] (1.7cm, 2.5cm) circle (.06cm);
        \draw (1.7cm, 2.44cm) -- (1.7cm, 2.0cm);
        \draw (1.5cm, 1.5cm) node [rotate=45] {JIT merge};  
      \end{tikzpicture}
    };
  \end{tikzpicture}
  \begin{itemize}
    \item first python example works!
      \begin{verbatim}
i = 0
while i < 1000000:
  i = i + 1
      \end{verbatim}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{June 2009}
  \begin{tikzpicture}[remember picture, overlay]
    \node [shift={(1cm, -4cm)}] at (current page.north west) {
      \begin{tikzpicture}[remember picture, overlay, font={\fontsize{8pt}{0pt} \ttfamily}]
        \node [shift={(-1cm, 1.4cm)}, above right] {
          \includegraphics[width=.15\textwidth]{pypy-logo.png}};
        \draw [thick, ->] (1.2cm, 2.5cm) -- (10cm, 2.5cm);
        \draw [fill=blue!20,very thin] (1.7cm, 2.5cm) circle (.06cm);
        \draw (1.7cm, 2.44cm) -- (1.7cm, 2.0cm);
        \draw (1.5cm, 1.5cm) node [rotate=45] {JIT merge};
        \draw [fill=blue!20,very thin] (2.3cm, 2.5cm) circle (.06cm);
        \draw (2.3cm, 2.44cm) -- (2.3cm, 2.0cm);
        \draw (1.95cm, 1.5cm) node [rotate=45] {first example};  
      \end{tikzpicture}
    };
  \end{tikzpicture}
  xxx
\end{frame}

\begin{frame}
  \frametitle{Nice graphs}
\end{frame}

\begin{frame}
  \frametitle{Status}
  \begin{block}{Status of today}
    \begin{itemize}
    \item JIT is fairly good
      \pause
    \item sometimes can consume a lot of memory
      \pause
    \item sometimes keep on compiling all the time
      \pause
    \item GCs are pretty good
      \pause
    \item VMs algorithms are not as polished as CPython
      ones [xxx reword]
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Near future}
  \begin{itemize}
    \item release March 2010
      \pause
    \item first PyPy release to contain JIT!
    \item mostly to try it out
  \end{itemize}
\end{frame}

\end{document}