### 2010-02-15 talk/pycon2010/pypyspeed/talk.tex
start working on talk

\documentclass[utf8x, 14pt]{beamer}

\mode<presentation>
{
  \usetheme{Warsaw}
}

\usepackage[english]{babel}

\usepackage[utf8x]{inputenc}
\usepackage{tikz}

\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{color}

\title{The speed of PyPy}

\author{Maciej Fijałkowski}

\institute[merlinux GmbH]
{ merlinux GmbH }

\date{Pycon 2010, February 20th 2010, Atlanta}

\begin{document}

\begin{frame}
  \titlepage
  \begin{figure}
    \begin{tabular}{c c c}
    \includegraphics[width=.30\textwidth]{../common/pypy-logo.png}
    &
    \hspace{2cm}
    &
    \includegraphics[width=.25\textwidth]{../common/merlinux-logo.png}
    \end{tabular}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{JIT - what's that about?}
  \begin{itemize}
    \item {\bf JIT is not a magical device!}
      \pause
    \item removes bytecode overhead
    \item removes frame overhead
    \item can make runtime decisions
    \item more classic optimization that can follow
\end{frame}

\begin{frame}
  \frametitle{The main idea}
  \begin{itemize}
    \item python has advanced features (frame introspection,
      arbitrary code execution, overloading globals)
    \item with JIT, you don't pay for them if you don't use
      them
    \item however, you pay if you use them, but they work
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{A piece of advice}
  \begin{itemize}
    \item Don't use advanced features if you don't have to
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{xxx}
\end{frame}

\end{document}