### 2010-02-17 talk/pycon2010/crossinterp/talk.tex
A rewrite of slides - needs more work

\documentclass[utf8x, 14pt]{beamer}

\mode<presentation>
{
  \usetheme{Warsaw}
}

\usepackage[english]{babel}

\usepackage[utf8x]{inputenc}
\usepackage{tikz}

\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{color}

\title{How to write cross-interpreter programs}

\author{Maciej Fijałkowski}

\institute[merlinux GmbH]
{ merlinux GmbH }

\date{Pycon 2010, February 19th 2010, Atlanta}

\begin{document}

\begin{frame}
  \titlepage
  \begin{figure}
    \begin{tabular}{c c c}
    \includegraphics[width=.20\textwidth]{../common/pypy-logo.png}
    &
    \hspace{2cm}
    &
    \includegraphics[width=.18\textwidth]{../common/merlinux-logo.png}
    \end{tabular}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{This talk}
  \begin{itemize}
    \item most people only target CPython (or Jython or IronPython)
    \item sometimes, you want your program to run on each of those
    \item libraries are more often cross-interpreter
      \pause
    \item hope to give you more reasons tomorrow
      \pause
    \item won't talk about py3k
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Have a way to verify compatibility}
  \begin{itemize}
    \item extensive test suite
    \item good coverage
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Exceptions}
  \begin{itemize}
    \item TypeError vs AttributeError change often between implementations,
      even CPython versions
    \item don't rely on exception string messages (they may differ)
    \item also means - don't use doctests
  \end{itemize}
  \begin{verbatim}
try:
  ...
except ImportError, ie:
  if str(ie) != '...':
    raise
  \end{verbatim}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Subclasses of builtin types and overriding}
  \begin{itemize}
    \item in general overridden methods on subclassed builtin types are not 
      invoked by preexisting other methods
      \begin{verbatim}
class d(dict):
  def __getitem__(self, e):
    ...
      \end{verbatim}
    \item would {\ttfamily keys()} go via this getitem?
    \item there can be corner cases, for example when
      there are more than one object involved
    \item tests are your friend
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Access to 3rdy party libraries}
  \begin{itemize}
    \item there is no good story here
    \item ctypes based access is going to be supported by all Pythons
    \item are there pure Python replacements/options?
    \item separate out dependencies/especially optional ones
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Don't rely on refcounting}
  \begin{itemize}
    \item example {\ttfamily open('x', 'w').write('stuff')}
    \item on refcounting, flushes file immediately
    \item on any other GC, it might be deferred for
      a while
    \item the single most-common problem when fixing twisted
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{\_\_del\_\_}
  \begin{itemize}
    \item in case of resurrection CPython will call {\ttfamily \_\_del\_\_} multiple times,
      other Pythons exactly once
    \item cycles with \_\_del\_\_s are not collected by CPython,
      PyPy breaks them randomly instead
    \item in PyPy (xxx Jython?) \_\_del\_\_ cannot be attached to classes after creation
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Use new-style classes}
  \begin{itemize}
    \item 3.x ready
    \item faster on PyPy, too
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{{\ttfamily sys.prefix}}
  \begin{itemize}
    \item implementations may have different installation layouts
    \item open issue, at least for PyPy, has compatibility consequences with
  setuptools
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{IO bytes vs unicode}
  \begin{itemize}
    \item convert/decode as soon as possible, keep text and bytes apart
    \item for 2.x Pythons use str for bytes and unicode for text
    \item the distinction is deeper in 3.x (str is unicode, bytes exist with
  slightly different interface than old str)
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Obscure corners}
  \begin{itemize}
    \item non-string keys in type dictionaries
    \item introspection results, implementation objects (e.g. builtin methods
      etc), may have different types
    \item exact naming of things (like list-comprehension variable)
  \end{itemize}
\end{frame}

\end{document}