### 2011-03-25 talk/icooolps2011/code/interpreter-slow.tex
- simplify the code a bit - delete the .py files now that the tex files are not generated any more

\begin{lstlisting}[mathescape,basicstyle=\ttfamily,numbers = right]
class Class(object):
    def __init__(self, name):
        self.name = name
        self.methods = {}

    def instantiate(self):
        return Instance(self)

    def find_method(self, name):
        return self.methods.get(name, None)

    def change_method(self, name, value):
        self.methods[name] = value

class Instance(object):
    def __init__(self, cls):
        self.cls = cls
        self.attributes = {}

    def getfield(self, name):
        return self.attributes.get(name, None)

    def write_attribute(self, name, value):
        self.attributes[name] = value

    def getattr(self, name):
	result = self.getfield(name)
	if result is None:
            result = self.cls.find_method(name)
	    if result is None:
		raise AttributeError
\end{lstlisting}