### 2011-03-25 talk/icooolps2011/code/map.tex

\begin{lstlisting}[mathescape,basicstyle=\ttfamily]
class Map(object):
    def __init__(self):
        self.indexes = {}
        self.other_maps = {}

    @purefunction
    def getindex(self, name):
        return self.indexes.get(name, -1)

    @purefunction
    def add_attribute(self, name):
        if name not in self.other_maps:
            newmap = Map()
            newmap.indexes.update(self.indexes)
            newmap.indexes[name] = len(self.indexes)
            self.other_maps[name] = newmap
        return self.other_maps[name]

EMPTY_MAP = Map()

class Instance(object):
    def __init__(self, cls):
        self.cls = cls
        self.map = EMPTY_MAP
        self.storage = []

    def getfield(self, name):
        map = hint(self.map, promote=True)
        index = map.getindex(name)
        if index != -1:
            return self.storage[index]
        return None

    def write_attribute(self, name, value):
        map = hint(self.map, promote=True)
        index = map.getindex(name)
        if index != -1:
            self.storage[index] = value
            return
        self.map = map.add_attribute(name)
        self.storage.append(value)

    def getattr(self, name):
	... # as before
\end{lstlisting}