### 2011-04-15 talk/icooolps2011/code/version.tex
some typos and problems I found while reading again

{\noop
\begin{lstlisting}[mathescape,basicstyle=\ttfamily]
class VersionTag(object):
    pass

class Class(object):
    def __init__(self, name):
        self.name = name
        self.methods = {}
        self.version = VersionTag()

    def find_method(self, name):
        promote(self)
        version = self.version
        promote(version)
        return self._find_method(name, version)

    @elidable
    def _find_method(self, name, version):
        assert version is self.version
        return self.methods.get(name, None)

    def write_method(self, name, value):
        self.methods[name] = value
        self.version = VersionTag()
\end{lstlisting}
}