### 2012-08-14 talk/vmil2012/paper.tex
Improve related work, evaluation and conclusion sections

\documentclass[10pt,preprint]{sigplanconf}

\usepackage{ifthen}
\usepackage{fancyvrb}
\usepackage{color}
\usepackage{wrapfig}
\usepackage{ulem}
\usepackage{xspace}
\usepackage{relsize}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[utf8]{inputenc}
\usepackage{setspace}
\usepackage[pdfpagelabels=true]{hyperref}
\usepackage[colorinlistoftodos]{todonotes}

\usepackage{listings}

\usepackage[T1]{fontenc}
\usepackage[scaled=0.81]{beramono}

\definecolor{commentgray}{rgb}{0.3,0.3,0.3}

\lstset{
  basicstyle=\ttfamily\footnotesize,
  language=Python,
  keywordstyle=\bfseries,
  stringstyle=\color{blue},
  commentstyle=\color{commentgray}\textit,
  fancyvrb=true,
  showstringspaces=false,
  %keywords={def,while,if,elif,return,class,get,set,new,guard_class}
  numberstyle = \tiny,
  numbersep = -20pt,
}

\hypersetup{%
  plainpages=false,%
  %hyperfootnotes=false,%
  colorlinks=true,%
  urlcolor=black,%
  citecolor=black,%
  linkcolor=black,%
  pdftitle={The Efficient Handling of Guards in the Design of RPython's Tracing JIT},%
  pdfauthor={David Schneider},
}

\newboolean{showcomments}
\setboolean{showcomments}{true}
\ifthenelse{\boolean{showcomments}}
  {\newcommand{\nb}[2]{
    \fbox{\bfseries\sffamily\scriptsize#1}
    {\sf\small$\blacktriangleright$\textit{#2}$\blacktriangleleft$}
   }
   \newcommand{\version}{\emph{\scriptsize$-$Id: main.tex 19055 2008-06-05 11:20:31Z cfbolz $-$}}
  }
  {\newcommand{\nb}[2]{}
   \newcommand{\version}{}
  }

\newcommand\cfbolz[1]{\nb{CFB}{#1}}
\newcommand\toon[1]{\nb{TOON}{#1}}
\newcommand\anto[1]{\nb{ANTO}{#1}}
\newcommand\arigo[1]{\nb{AR}{#1}}
\newcommand\fijal[1]{\nb{FIJAL}{#1}}
\newcommand\pedronis[1]{\nb{PEDRONIS}{#1}}
\newcommand\bivab[1]{\nb{DAVID}{#1}}
\newcommand{\commentout}[1]{}

\newcommand{\noop}{}

\newcommand\ie{i.e.,\xspace}
\newcommand\eg{e.g.,\xspace}

\normalem

\let\oldcite=\cite

\renewcommand\cite[1]{\ifthenelse{\equal{#1}{XXX}}{[citation~needed]}{\oldcite{#1}}}

\definecolor{gray}{rgb}{0.5,0.5,0.5}

\begin{document}

\title{The Efficient Handling of Guards in the Design of RPython's Tracing JIT}

\authorinfo{David Schneider$^{a}$ \and Carl Friedrich Bolz$^a$}
           {$^a$Heinrich-Heine-Universität Düsseldorf, STUPS Group, Germany
           }
           {david.schneider@uni-duesseldorf.de \and cfbolz@gmx.de}

\conferenceinfo{VMIL'12}{}
\CopyrightYear{2012}
\crdata{}

\maketitle

\category{D.3.4}{Programming Languages}{Processors}[code generation,
incremental compilers, interpreters, run-time environments]

\terms
Languages, Performance, Experimentation

\keywords{tracing JIT, guards, deoptimization}

\begin{abstract}
Tracing just-in-time (JIT) compilers record linear control flow paths,
inserting operations called guards at points of possible divergence. These
operations occur frequently generated traces and therefore it is important to
design and implement them carefully to find the right trade-off between
execution speed, deoptimization,
and memory overhead.  In this paper we describe the design decisions about
guards taken in the implementation of the RPython tracing JIT. Furthermore we
measure various properties of guards.
% \o/
\end{abstract}

%___________________________________________________________________________
\todo{find a better name for \texttt{low-level resume data}}
\todo{find better names for JIT front- and backend}
\todo{mention somewhere that it is to be expected that most guards do not fail}
\section{Introduction}

\todo{the introduction needs some work}
\cfbolz{the first two two paragraphs talk about deoptimization, then it
switches to guards. I would say we should only talk about guards in the
beginning}
Tracing just-in-time (JIT) compilers record and compile linear control flow paths of operations executed by an interpreter
inserting operations called guards at points of possible divergence.
In this paper we describe and analyze how guards work, explaining what concepts
are used in the intermediate and low-level representation of the JIT
instructions and how these are implemented.

Our aim is to help understand the design constraints when
implementing guards. Guards have a runtime cost, they take time to execute. On
the other hand, guards are possible deoptimization points, meaning the recorded
and compiled path has to be left returning control to the interpreter. They need
enough associated information to enable rebuilding the interpreter state.
Based on the informal observation that guards are among the most common
operations in the traces produced by RPython's tracing JIT, our
goal is to present concrete numbers for the frequency and the overhead related
to guards, explain how they are implemented in the different levels of this particular
tracing JIT and clarify the rationale behind the design decisions based on the
numbers provided here.

The operations executed by an interpreter are recorded by the tracing JIT in
case they are frequently executed (this process is described in more detail in
Section \ref{sec:Resume Data}). During the recording phase \texttt{guards} are
inserted into the recorded trace at all
points where the control flow could diverge. As can be seen in
Figure~\ref{fig:guard_percent} guards account for about 14\% to 22\% of the
operations before and for about 15\% to 20\% of the operations after
optimizing the traces generated for the different benchmarks used in this paper.
This makes guards one of the most common types of operations. As this paper will show, many of these guards
fail rarely or not all during execution.

There are several aspects to consider
in the design and optimization of guards, the first aspect is that due to the
large number of guards the memory overhead related to storing the information
needed for deoptimization should be kept low. A second aspect is that
successfully checking guards, i.e. not leaving the compiled trace, – which is
the common case – should be a cheap operation to execute favouring the on-trace
execution speed in contrast to the deoptimization case where the state has to
be rebuilt using the stored information. These constraints and trade-offs are
what make the design and optimization of guards an important and non-trivial
aspect of the low-level design of a tracing just-in-time compiler.

%Section~\ref{sec:Evaluation} presents Figures about the absolute number of
%operations for each benchmark, and the overhead produced by the information
%stored at the different levels for the guards
In this paper we want to substantiate the aforementioned observations and
describe based on them the reasoning behind the implementation of guards in
RPython's tracing just-in-time compiler. the contributions of this paper are:
\begin{itemize}
  \item An analysis and benchmark of guards in the context of RPython's tracing JIT.
  %An analysis of guards in the context of RPython's tracing JIT to
  %substantiate the aforementioned observation, based on a set of benchmarks,
  \item detailed measurements about the frequency and the
  overhead associated with guards, and
  \item a description about how guards are implemented in the high\-
  and low-level components of the JIT and describe the rationale behind the design.
\end{itemize}

\begin{figure}
    \include{figures/guard_table}
    \caption{Percentage of guards before and after optimization for different benchmarks}
    \label{fig:guard_percent}
\end{figure}

The set of central concepts upon which this work is based is described in
Section~\ref{sec:Background}, such as the PyPy project, the RPython language
and its meta-tracing JIT. Based on these concepts in Section~\ref{sec:Resume
Data} we proceed to describe for RPython's tracing JIT the details of guards in
the frontend\bivab{better term for this?} related to recording and storing the
information required to rebuild the interpreter state in case of a guard
failure. Once the frontend has traced and optimized a loop it invokes the
backend to compile the operations to machine code, Section \ref{sec:Guards in
the Backend} describes the low-level aspects of how guards are implemented in
the JIT-backend. The frequency of guards and the overhead associated with the
implementation described in this paper is discussed in
Section~\ref{sec:evaluation}. Section~\ref{sec:Related Work} presents an
overview about how guards are treated in the context of other just-in-time
compilers. Finally Section~\ref{sec:Conclusion} summarizes our conclusions and
gives an outlook on further research topics.

\section{Background}
\label{sec:Background}

\subsection{RPython and the PyPy Project}
\label{sub:pypy}

The RPython language and the PyPy project were started in 2002 with the goal of
creating a Python interpreter written in a high level language, allowing easy
language experimentation and extension. PyPy is now a fully compatible
alternative implementation of the Python language, that is on average about 5 times faster than the reference implementation. The
implementation takes advantage of the language features provided by RPython
such as the provided tracing just-in-time compiler described below.

RPython, the language and the toolset originally developed to implement the
Python interpreter have developed into a general environment for experimenting
and developing fast and maintainable dynamic language implementations. There
are, besides the Python interpreter, implementations of Prolog, Javascript, R,
Smalltalk among other that are written in RPython at different levels of
completeness.

RPython is constructed from two components:
\begin{itemize}
 \item the language itself
 \item the translation toolchain used to transform RPython programs to executable units
\end{itemize}
The RPython language
is a statically typed object oriented high level language. The language provides
several features such as automatic memory management
and just-in-time compilation. When writing an interpreter using RPython the
programmer only has to write the interpreter for the language she is
implementing.  The second RPython component, the translation toolchain, is used
to transform the program to a low level representations suited to be compiled
and run on one of the different supported target platforms/architectures such
as C, .NET and Java. During the transformation process
different low level aspects suited for the target environment are automatically
added to the program such as (if needed) a garbage collector
and based on hints provided by the author a just-in-time compiler.

\subsection{RPython's Tracing JIT Compilers}
\label{sub:tracing}
Tracing is a technique of just-in-time compilers that generate code by
observing the execution of a program. VMs using tracing JITs are typically
mixed-mode execution environments that also contain an interpreter. The
interpreter profiles the executing program and selects frequently executed code
paths to be compiled to machine code. Many tracing JIT compilers focus on
selecting hot loops.

After profiling identifies an interesting
path, tracing is started thus recording all operations that are executed on this
path. This includes inlining functional calls.
As in most compilers, tracing JITs use an intermediate representation to
store the recorded operations, which is typically in SSA
form~\cite{cytron_efficiently_1991}. Since tracing follows actual execution the
code that is recorded
represents only one possible path through the control flow graph. Points of
divergence from the recorded path are marked with special operations called
\emph{guards}, these operations ensure that assumptions valid during the
tracing phase are still valid when the code has been compiled and is executed.
In the case of dynamic languages, guards are also used to encode type checks
that come from optimistic type specialization by recording the types of
variables seen during tracing.
After a trace has been recorded it is optimized and then compiled to platform
specific machine code.

When the check of a guard fails, the execution of the machine code must be
stopped and the control is returned to the interpreter, after the interpreter's
state has been restored. If a particular guard fails often a new trace is
recorded starting from the guard. We will refer to this kind of trace as a
\emph{bridge}. Once a bridge has been traced it is attached to the
corresponding guard by patching the machine code. The next time the guard fails
the bridge will be executed instead of leaving the machine code.

RPython provides a tracing JIT that can be reused for a number of language
implementations. This is possible, because it traces the execution of the
language interpreter instead of tracing the user program directly. This
approach is called \emph{meta-tracing}. For the purpose of this paper the fact
that RPython's tracing JIT is a meta-tracing JIT can be ignored.

\begin{figure}
    \input{figures/example.tex}
    \caption{Example Program}
    \label{fig:example}
\end{figure}

Figure~\ref{fig:example} shows an example RPython function that checks
whether a number reduces to 1 with less than 100 steps of the Collatz process.\footnote{\url{http://en.wikipedia.org/wiki/Collatz_conjecture}}
It uses an \lstinline{Even} and an \lstinline{Odd} class to box the numbers, to
make the example more interesting. If the loop in \lstinline{check_reduces} is
traced when \lstinline{a} is a multiple of four, the unoptimized
trace looks like in Figure~\ref{fig:unopt-trace}. The line numbers in the trace
correspond to the line numbers in Figure~\ref{fig:trace-log}. The resulting
trace repeatedly halves the current value and checks whether it is equal to
one, or odd. In either of these cases the trace is left via a guard failure.

\begin{figure}
    \input{figures/unopt-log.tex}
    \caption{Unoptimized trace, the line numbers in the trace correspond to the line numbers in Figure~\ref{fig:trace-log}.}
    \label{fig:unopt-trace}
\end{figure}

\section{Guards in the Frontend} %{Resume Data}
\label{sec:Resume Data}

Since tracing linearizes control flow by following one concrete execution,
the full control flow of a program is not observed.
The possible points of deviation from the trace are denoted by guard operations
that check whether the same assumptions observed while tracing
still hold during execution.
Similarly, in the case of dynamic languages guards can also encode type
assumptions.
In later executions of the trace the guards can fail.
If that happens, execution needs to continue in the interpreter.
This means it is necessary to attach enough information to a guard
to reconstruct the interpreter state when that guard fails.
This information is called the \emph{resume data}.

To do this reconstruction it is necessary to take the values of the SSA
variables of the trace and build interpreter stack frames.  Tracing
aggressively inlines functions, therefore the reconstructed state of the
interpreter can consist of several interpreter frames.

If a guard fails often enough, a trace is started from it
forming a trace tree.
When that happens another use case of resume data
is to construct the tracer state.
After the bridge has been recorded and compiled it is attached to the guard.
If the guard fails later the bridge is executed. Therefore the resume data of
that guard is no longer needed.

There are several forces guiding the design of resume data handling.
Guards are a very common operations in the traces.
However, a large percentage of all operations
are optimized away before code generation.
Since there are a lot of guards
the resume data needs to be stored in a very compact way.
On the other hand, tracing should be as fast as possible,
so the construction of resume data must not take too much time.

\subsection{Capturing of Resume Data During Tracing}
\label{sub:capturing}

Every time a guard is recorded during tracing
the tracer attaches preliminary resume data to it.
The data is preliminary in that it is not particularly compact yet.
The preliminary resume data takes the form of a stack of symbolic frames.
The stack contains only those interpreter frames seen by the tracer.
The frames are symbolic in that the local variables in the frames
do not contain values.
Instead, every local variables contains the SSA variable of the trace
where the value would later come from, or a constant.

\subsection{Compression of Resume Data}
\label{sub:compression}

After tracing has been finished the trace is optimized.
During optimization a large percentage of operations can be removed.
In the process the resume data is transformed into its final, compressed form.
The rationale for not compressing the resume data during tracing
is that a lot of guards will be optimized away.
For them, the compression effort would be lost.

The core idea of storing resume data as compactly as possible
is to share parts of the data structure between subsequent guards.
This is useful because the density of guards in traces is so high,
that quite often not much changes between them.
Since resume data is a linked list of symbolic frames
in many cases only the information in the top frame changes from one guard to the next.
The other symbolic frames can often just be reused.
The reason for this is that during tracing only the variables
of the currently executing frame can change.
Therefore if two guards are generated from code in the same function
the resume data of the rest of the frame stack can be reused.

In addition to sharing as much as possible between subsequent guards
a compact representation of the local variables of symbolic frames is used.
Every variable in the symbolic frame is encoded using two bytes.
Two bits are used as a tag to denote where the value of the variable
comes from.
The remaining 14 bits are a payload that depends on the tag bits.

The possible source of information are:

\begin{itemize}
    \item For small integer constants
        the payload contains the value of the constant.
    \item For other constants
        the payload contains an index into a per-loop list of constants.
    \item For SSA variables,
        the payload is the number of the variable.
    \item For virtuals,
        the payload is an index into a list of virtuals, see next section.
\end{itemize}

\subsection{Interaction With Optimization}
\label{sub:optimization}

Guards interact with optimizations in various ways.
Using many classical compiler optimizations the JIT tries to remove as many
operations, and therefore guards, as possible.
In particular guards can be removed by subexpression elimination.
If the same guard is encountered a second time in the trace,
the second one can be removed.
This also works if a later guard is weaker
and hence implied by an earlier guard.

One of the techniques in the optimizer specific to tracing for removing guards
is guard strengthening~\cite{bebenita_spur:_2010}.
The idea of guard strengthening is that if a later guard is stronger
than an earlier guard it makes sense to move the stronger guard
to the point of the earlier, weaker guard and to remove the weaker guard.
Moving a guard to an earlier point is always valid,
it just means that the guard fails earlier during the trace execution
(the other direction is clearly not valid).

The other important point of interaction between resume data and the optimizer
is RPython's allocation removal optimization~\cite{bolz_allocation_2011}.
This optimization discovers allocations in the trace that create objects
that do not survive long.
An example is the instance of \lstinline{Even} in Figure~\ref{fig:unopt-trace}.
Allocation removal makes resume data more complex.
Since allocations are removed from the trace it becomes necessary
to reconstruct the objects that were not allocated so far when a guard fails.
Consequently the resume data needs to store enough information
to make this reconstruction possible.

Adding this additional information is done as follows:
So far, every variable in the symbolic frames
contains a constant or an SSA variable.
After allocation removal the variables in the symbolic frames can also contain
``virtual'' objects.
These are objects that were not allocated so far,
because the optimizer removed their allocation.
The structure of the heap objects that have to be allocated on guard failure
is described by the virtual objects stored in the symbolic frames.
To this end, the content of every field of the virtual object is described
in the same way that the local variables of symbolic frames are described.
The fields of the virtual objects can therefore be SSA variables, constants
or other virtual objects.
They are encoded using the same compact two-byte representation
as local variables.

During the storing of resume data virtual objects are also shared
between subsequent guards as much as possible.
The same observation as about frames applies:
Quite often a virtual object does not change from one guard to the next.
Then the data structure is shared.

A related optimization is the handling of heap stores by the optimizer.
The optimizer tries to delay stores into the heap as long as possible.
This is done because often heap stores become unnecessary
due to another store to the same memory location later in the trace.
This can make it necessary to perform these delayed stores
when leaving the trace via a guard.
Therefore the resume data needs to contain a description
of the delayed stores to be able to perform them when the guard fails.
So far no special compression is done with this information,
compared to the other source of information delayed heap stores are quite rare.

Figure~\ref{fig:trace-log} shows the optimized version of the trace in
Figure~\ref{fig:unopt-trace}. Allocation removal has removed the
\lstinline{new} operation and other operations handling the instance. The
operations handle unboxed numbers now.

Figure~\ref{fig:resume-data} sketches the symbolic frames of the first two
guards in the trace. The frames for \lstinline{check_reduces} and
\lstinline{Even.step} as well as the description of the allocation-removed
virtual instance of \lstinline{Even} are shared between the two guards.

\todo{fix labels in diagram}

% section Resume Data (end)

\begin{figure}
    \input{figures/log.tex}
    \caption{Optimized trace}
    \label{fig:trace-log}
\end{figure}
% section Resume Data (end)
\section{Guards in the Backend}
\label{sec:Guards in the Backend}

\begin{figure}
\includegraphics[width=0.5\textwidth]{figures/resume_data.pdf}
\caption{The resume data for Figure~\ref{fig:trace-log}}
\label{fig:resume-data}
\end{figure}

After optimization the resulting trace is handed over to the platform specific
backend to be compiled to machine code. The compilation phase consists of two
passes over the lists of instructions, a backwards pass to calculate live
ranges of IR-level variables and a forward pass to emit the instructions. During
the forward pass IR-level variables are assigned to registers and stack
locations by the register allocator according to the requirements of the
emitted instructions.  Eviction/spilling is performed based on the live range
information collected in the first pass. Each IR instruction is transformed
into one or more machine level instructions that implement the required
semantics. Operations without side effects whose result is not used are not
emitted. Guards instructions are transformed into fast checks at the machine
code level that verify the corresponding condition.  In cases the value being
checked by the guard is not used anywhere else the guard and the operation
producing the value can often be merged, further reducing the overhead of the guard.
Figure \ref{fig:trace-compiled} shows how the \texttt{int\_eq} operation
followed by a \texttt{guard\_false} from the trace in Figure~\ref{fig:trace-log} are compiled to
pseudo-assembler if the operation and the guard are compiled separated or if
they are merged.

\todo{Figure needs better formatting}
\begin{figure}[ht]
  \noindent
  \centering
  \begin{minipage}{1\columnwidth}
\begin{lstlisting}[mathescape, numbers=right, escapechar=|, firstnumber=18]
$b_3$ = int_eq($i_5$, 1)                    |\setcounter{lstnumber}{17}|
guard_false($b_3$)                          |\setcounter{lstnumber}{-1}|
\end{lstlisting}
  \end{minipage}
  \begin{minipage}{.40\columnwidth}
    \begin{lstlisting}
CMP r6, #1
MOVEQ r8, #1
MOVNE r8, #0
CMP r8, #0
BEQ <bailout>
    \end{lstlisting}
  \end{minipage}
  \hfill
  \begin{minipage}{.40\columnwidth}
    \begin{lstlisting}
CMP r6, #1
BNE <bailout>
...
...
...
    \end{lstlisting}
  \end{minipage}
  \caption{Result of separated (left) and merged (right) compilation of operations and guards (top).}
  \label{fig:trace-compiled}
\end{figure}

Attached to each guard in the IR is a list of the IR-variables required to
rebuild the execution state in case the trace is left through
the guard. When a guard is compiled, in addition to the
condition check two things are generated/compiled.

First a special data
structure called \emph{low-level resume data} is created. This data structure encodes the
information about where, i.e. which register or stack location, the IR-variables required to rebuild the state will be stored when the guard is executed.
This data
structure stores the values in a succinct manner using an encoding that uses
8 bits to store 7 bits of information, ignoring leading zeros. This encoding is efficient to create and
provides a compact representation of the needed information in order
to maintain an acceptable memory profile.

Second for each guard a piece of code is generated that acts as a trampoline.
Guards are implemented as a conditional jump to this trampoline in case the
guard check fails.
In the trampoline the pointer to the
\emph{low-level resume data} is loaded and after storing the current execution state
(registers and stack) execution jumps to a generic bailout handler, also known
as \texttt{compensation code},
that is used to leave the compiled trace in case of a guard failure.

Using the encoded location information the bailout handler reads from the
saved execution state the values that the IR-variables had  at the time of the
guard failure and stores them in a location that can be read by the frontend.
After saving the information the control is passed to the frontend signaling
which guard failed so the frontend can read the information passed and restore
the state corresponding to the point in the program.

As in previous sections the underlying idea for the design of guards is to have
a fast on-trace profile and a potentially slow one in the bailout case where
the execution has to return to the interpreter due to a guard failure. At the same
time the data stored in the backend, required to rebuild the state, should be as
compact as possible to reduce the memory overhead produced by the large number
of guards, the numbers in Figure~\ref{fig:backend_data} illustrate that the
compressed encoding currently has about 15\% to 25\% of the size of of the
generated instructions on x86.

As explained in previous sections, when a specific guard has failed often enough
a new trace, referred to as a \emph{bridge}, starting from this guard is recorded and
compiled.
Since the goal of compiling bridges is to improve execution speed on the
diverged path (failing guard) they should not introduce additional overhead.
In particular the failure of the guard should not lead
to leaving the compiled code prior to execution the code of the bridge.

The process of compiling a bridge is very similar to compiling a loop.
Instructions and guards are processed in the same way as described above. The
main difference is the setup phase. When compiling a trace we start with a clean
slate. The compilation of a bridge is started from a state (register and stack
bindings) that corresponds to the state during the compilation of the original
guard. To restore the state needed to compile the bridge we use the encoded
representation created for the guard to rebuild the bindings from IR-variables
to stack locations and registers used in the register allocator.  With this
reconstruction all bindings are restored to the state as they were in the
original loop up to the guard. This means that no register/stack reshuffling is
needed before executing a bridge.

Once the bridge has been compiled the corresponding guard is
patched to redirect control flow to the bridge in case the check fails. In
the future, if the guard fails again it jumps to the code compiled for the bridge
instead of bailing out. Once the guard has been compiled and attached to the
loop the guard becomes just a point where control-flow can split. The loop
after the guard and the bridge are just conditional paths.
Figure~\ref{fig:trampoline} shows a diagram of a compiled loop with two guards,
Guard \#1 jumps to the trampoline, loads the \texttt{low level resume data} and
then calls the bailout handler, whereas Guard \#2 has already been patched
and directly jumps to the corresponding bridge. The bridge also contains two
guards that work based on the same principles.
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/loop_bridge.pdf}
\caption{Trace control flow in case of guard failures with and without bridges}
\label{fig:trampoline}
\end{figure}
% section Guards in the Backend (end)

%___________________________________________________________________________

\section{Evaluation}
\label{sec:evaluation}
\todo{improve the table formatting}

The results presented in this section are based on numbers gathered by running
a subset of the standard PyPy benchmarks. The PyPy benchmarks are used to
measure the performance of PyPy and are composed of a series of
micro-benchmarks and larger programs.\footnote{\url{http://speed.pypy.org/}} The
benchmarks were taken from the PyPy benchmarks repository using revision
\texttt{ff7b35837d0f}.\footnote{\url{https://bitbucket.org/pypy/benchmarks/src/ff7b35837d0f}}
The benchmarks were run on a version of PyPy based on the
tag~\texttt{0b77afaafdd0} and patched to collect additional data about the
guards in the machine code
backends.\footnote{\url{https://bitbucket.org/pypy/pypy/src/0b77afaafdd0}} The
tools used to run and evaluate the benchmarks including the patches applied to
the PyPy sourcecode can be found in the repository for this
paper.\footnote{\url{https://bitbucket.org/pypy/extradoc/src/tip/talk/vmil2012}}
All
benchmark data was collected on a MacBook Pro 64 bit running Max OS 10.8 with
the loop unrolling optimization disabled.\footnote{Since loop unrolling
duplicates the body of loops it would no longer be possible to meaningfully
compare the number of operations before and after optimization. Loop unrolling
is most effective for numeric kernels, so the benchmarks presented here are not
affected much by its absence.}

We used the following benchmarks:

\begin{description}
    \item[chaos:] A Chaosgame implementation creating a fractal.
    \item[crypto\_pyaes:] An AES implementation.
    \item[django:] The templating engine of the Django Web
        framework.\footnote{\url{http://www.djangoproject.com/}}

    \item[go:] A Monte-Carlo Go
        AI.\footnote{\url{http://shed-skin.blogspot.com/2009/07/disco-elegant-python-go-player.html}}
    \item[pyflate\_fast:] A BZ2 decoder.
    \item[raytrace\_simple:] A ray tracer.
    \item[richards:] The Richards benchmark.
    \item[spambayes:] A Bayesian spam filter.\footnote{\url{http://spambayes.sourceforge.net/}}
    \item[simpy\_expand:] A computer algebra system.
    \item[telco:] A Python version of the Telco decimal
        benchmark,\footnote{\url{http://speleotrove.com/decimal/telco.html}}
        using a pure Python decimal floating point implementation.
    \item[twisted\_names:] A DNS server benchmark using the Twisted networking
        framework.\footnote{\url{http://twistedmatrix.com/}}
\end{description}

From the mentioned benchmarks we collected different datasets to evaluate the
frequency, the overhead and overall behaviour of guards, the results are
summarized in the remainder of this section. We want to point out three
aspects of guards in particular:
\begin{itemize}
  \item Guards are very common operations in traces.
  \item There is overhead associated with guards.
  \item Guard failures are local and rare.
\end{itemize}

All measurements presented in this section do not take garbage collection of machine code into account. Pieces
of machine code can be globally invalidated or just become cold again. In both
cases the generated machine code and the related data is garbage collected. The
figures show the total amount of operations that are evaluated by the JIT and
the total amount of code and data that is generated from the optimized traces.

\subsection{Frequency of Guards}
\label{sub:guard_frequency}
\begin{figure*}
    \include{figures/benchmarks_table}
    \caption{Number of operations in the recorded traces and the relative amount of guards before and after optimizations}
    \label{fig:benchmarks}
\end{figure*}

Figure~\ref{fig:benchmarks} extends Figure~\ref{fig:guard_percent} and summarizes the total number of operations that were
recorded during tracing for each of the benchmarks and what percentage of these
operations are guards. The number of operations was counted on the unoptimized
and optimized traces. The Figure shows that the overall optimization rate for
operations, which is between 69.4\% and 83.89\%, of the traced operations and the
optimization rate of guards, which is between 65.8\% and 86.2\% of the
operations, are very similar. This indicates that the optimizer can remove
most of the guards, but after the optimization pass these still account for
15.2\% to 20.2\% of the operations being compiled and later executed.
The frequency of guard operations makes it important to store the associated
information efficiently and also to make sure that guard checks are executed
quickly.

\subsection{Overhead of Guards}
\label{sub:guard_overhead}
\begin{figure}
    \include{figures/resume_data_table}
    \caption{Resume data sizes}
    \label{fig:resume_data_sizes}
\end{figure}

The overhead that is incurred by the JIT to manage the \texttt{resume data},
the \texttt{low-level resume data} as well as the generated machine code is
shown in Figure~\ref{fig:backend_data}. It shows the total memory consumption
of the code and of the data generated by the machine code backend and an
approximation of the size of the \texttt{resume data} structures for the
different benchmarks mentioned above. The machine code taken into account is
composed of the compiled operations, the trampolines generated for the guards
and a set of support functions that are generated when the JIT starts and which
are shared by all compiled traces. The size of the \texttt{low-level resume
data} is the size of the compressed mapping from registers and stack to
IR-level variables and finally the size of the \texttt{resume data} is an
approximation of the size of the compressed high-level resume data as described
in Section~\ref{sec:Resume Data}.\footnote{
The size of the resume data is not measured at runtime, but reconstructed from
log files.}

For the different benchmarks the \texttt{low-level resume data} has a size of
about 15\% to 20\% of the amount of memory compared to the size of the
generated machine code. On the other hand the generated machine code has only a
size ranging from 20.5\% to 37.98\% of the size of the high and low-level
resume data combined and being compressed as described before.

Tracing JIT compilers only compile the subset of the code executed in a program
that occurs in a hot loop, for this reason the amount of generated machine
code will be smaller than in other juts-in-time compilation approaches.  This
creates a larger discrepancy between the size of the \texttt{resume data} when
compared to the size of the generated machine code and illustrates why it is important to compress the \texttt{resume data} information.

\begin{figure}
    \include{figures/backend_table}
    \caption{Total size of generated machine code and resume data}
    \label{fig:backend_data}
\end{figure}

Why the efficient storing of the \texttt{resume data} is a central concern in the design
of guards is illustrated by Figure~\ref{fig:resume_data_sizes}. This figure shows
the size of the compressed \texttt{resume data}, the approximated size of
storing the \texttt{resume data} without compression and
an approximation of the best possible compression of the resume data by
compressing the data using the
\texttt{xz} compression tool, which is a ``general-purpose data compression
software with high compression ratio''.\footnote{\url{http://tukaani.org/xz/}}

The results show that the current approach of compression and data sharing only
requires 18.3\% to 31.1\% of the space compared to a naive approach. This
shows that large parts of the resume data are redundant and can be stored more
efficiently using the techniques described earlier. On the other hand
comparing the results to the xz compression which only needs between 17.1\%
and 21.1\% of the space required by our compression shows that the compression
is not optimal but a trade-off between the required space and the time needed
to build a good, compressed representation of the resume data for the
large amount of guards present in the traces.

\subsection{Guard Failures}
\label{sub:guard_failure}
The last point in this discussion is the frequency of guard failures.
Figure~\ref{fig:failing_guards} presents for each benchmark a list of the
relative amounts of guards that ever fail and of guards that fail often enough that a bridge is compiled.
\footnote{
    The threshold used is 200 failures. This rather high threshold was picked experimentally to give
    good results for long-running programs.
}

The numbers presented for guards that have a bridge represent the
failures up to the compilation of the bridge and all executions of the then
attached bridge.

\begin{figure}
    \include{figures/failing_guards_table}
    \caption{Failing guards relative to the total number of guards}
    \label{fig:failing_guards}
\end{figure}

From Figure~\ref{fig:failing_guards} we can see that only a very small amount
of all the guards in the optimized traces ever fail. This amount varies between
2.4\% and 5.7\% of all guards. As can be expected, even fewer guards fail often
enough that a bridge is compiled for them, only 1.2\% to 3.6\% of all guards
fail often enough that a bridge is compiled. Also, of all failing guards a few fail extremely often
and most fail rarely. The results emphasize that as most of the guards never
fail it is important to make sure that the successful execution of a guard does
not have unnecessary overhead.

\section{Related Work}
\label{sec:Related Work}

\subsection{Guards in Other Tracing JITs}
\label{sub:Guards in Other Tracing JITs}

Guards, as described, are a concept associated with tracing just-in-time
compilers to represent possible divergent control flow paths.

SPUR~\cite{bebenita_spur:_2010} is a tracing JIT compiler
for a C\# virtual machine.
It handles guards by always generating code for every one of them
that transfers control back to the unoptimized code.
Since the transfer code needs to reconstruct the stack frames
of the unoptimized code,
the transfer code is quite large.

Mike Pall, the author of LuaJIT describes in a post to the lua-users mailing
list different technologies and techniques used in the implementation of
LuaJIT~\cite{Pall:2009}. Pall explains that guards in LuaJIT use a datastucture
called snapshots, similar to RPython's resume data, to store the information
about how to rebuild the state from a guard failure using the information in the
snapshot and the machine execution state. According to Pall~\cite{Pall:2009}
snapshots for guards in LuaJIT are associated with a large memory footprint.
The solution used there is to store sparse snapshots, avoiding the creation
of snapshots for every guard to reduce memory pressure. Snapshots are only
created for guards after updates to the global state, after control flow points
from the original program and for guards that are likely to fail. As an outlook
Pall mentions the plans to switch to compressed snapshots to further reduce
redundancy. The approach of not creating snapshots at all for every guard is
orthogonal to the resume data compression presented in this paper and could be
reused within RPython to improve the memory usage further.

Linking side exits to pieces of later compiled machine code was described first
in the context of Dynamo~\cite{Bala:2000wv} under the name of Fragment Linking.
Once a new hot trace is emitted into the fragment cache it is linked to side
exit that led to the compilation of the fragment. Fragment Linking avoids the
performance penalty involved in leaving the compiled code. Fragment linking
also allows to remove compensation code associated to the linked fragments that
would have been required to restored the execution state on the side exit.

Gal et. al~\cite{Gal:2006} describe how in the HotpathVM they experimented
with having one generic compensation code block, like the RPython JIT, that
uses a register variable mapping to restore the interpreter state. Later this
was replaced by generating compensation code for each guard which produced a
lower overhead in their benchmarks. HotpathVM also records secondary traces
starting from failing guards that are connected directly to the original trace.
Secondary traces are compiled by first restoring the register allocator state to
the state at the side exit. The information is retrieved from a mapping stored
in the guard that maps machine level registers and stack to Java level stack
and variables.

TraceMonkey, a tracing JIT for JavaScript, uses trace stitching
to avoid the overhead of returning to the trace monitor and calling another
trace when taking a side exit~\cite{Gal:2009ux}. In this approach it is required to write live
values to an activation record before entering the new trace.

% subsection Guards in Other Tracing JITs (end)

\subsection{Deoptimization in Method-Based JITs}
\label{sub:Deoptimization in Method-Based JITs}

Deoptimization in method-based JITs is used if one of the assumptions
of the code generated by a JIT-compiler changes.
This is often the case when new code is added to the system,
or when the programmer tries to debug the program.

Deutsch et. al.~\cite{deutsch_efficient_1984} describe the use of stack descriptions
to make it possible to do source-level debugging of JIT-compiled code.
Self uses deoptimization to reach the same goal~\cite{XXX}.
When a function is to be debugged, the optimized code version is left
and one compiled without inlining and other optimizations is entered.
Self uses scope descriptors to describe the frames
that need to be re-created when leaving the optimized code.
The scope descriptors are between 0.45 and 0.76 times
the size of the generated machine code.

Java Hotspot~\cite{paleczny_java_2001} contains a deoptimization framework that is used
for debugging and when an uncommon trap is triggered.
To be able to do this, Hotspot stores a mapping from optimized states
back to the interpreter state at various deoptimization points.
There is no discussion of the memory use of this information.

The deoptimization information of Hotspot is extended
to support correct behaviour
when scalar replacement of fields is done for non-escaping objects~\cite{kotzmann_escape_2005}.
The approach is extremely similar to how RPython's JIT handles virtual objects.
For every object that is not allocated in the code,
the deoptimization information contains a description
of the content of the fields.
When deoptimizing code, these objects are reallocated
and their fields filled with the values
described by the deoptimization information.
The paper does not describe any attempts to store this information compactly.
This may not be needed in their approach, because method-based JITs have
fewer deoptimization points than tracing JITs.

% subsection Deoptimization in Method-Based JITs (end)
% section Related Work (end)

\section{Conclusion}
\label{sec:Conclusion}
In this paper we have concentrated on guards, an operation found in
tracing just-in-time compilers and used to denote points of possible control
flow divergence in recorded traces.
Based on the observation that guards are a frequent operation in traces and
that they do not fail often, we described how they have been implemented in the
high and low level components of RPython's tracing JIT compiler.

Additionally we have presented experimental data collected using the standard PyPy
benchmark set to evaluate previous observations and assumptions. Our
experiments confirmed that guards are a very common
operation in traces. At the same time guards are associated with a high
overhead, because for all compiled guards information needs to be
stored to restore the execution state in case of a bailout. The measurements
showed that the compression techniques used in PyPy effectively reduce the
overhead of guards, but it still produces a significant overhead. The results
also showed that guard failure is a local event: there are few
guards that fail at all, and even fewer that fail very often.
These numbers validate the design decision of reducing the overhead of
successful guard checks as much as possible while paying a higher price in the
case of bailout due to having to decode a compressed state representation.
The compressed state representation reduces the memory footprint of rarely
used data.

Based on the observation that guard failure is rare it
would be worth exploring if a more aggressive compression scheme for guards
would be worth the memory saving in contrast to the increased decoding
overhead. Based on the same observation we would like to explore the concept of
LuaJIT's sparse snapshots and its applicability to PyPy.

\section*{Acknowledgements}
We would like to thank David Edelsohn and Stephan Zalewski for their helpful
feedback and valuable comments while writing this paper.

%\section*{Appendix}
%\todo{remove this section and the figures}
%\begin{figure*}
%    \include{figures/ops_count_table}
%    \caption{Relative numbers of operations in the traces generated for
%    different benchmarks}
%    \label{fig:ops_count}
%\end{figure*}
%\begin{figure*}
%\centering
%\includegraphics[width=\textwidth]{figures/ops_pie.pdf}
%\caption{Relative frequency of operations before and after optimization}
%\label{fig:ops_pie}
%\end{figure*}
\bibliographystyle{abbrv}
\bibliography{zotero,paper}
\listoftodos
\end{document}