### 2012-10-18 talk/vmil2012/presentation/talk.tex
interleave example with explanation slides

\documentclass[utf8x]{beamer}

% This file is a solution template for:

% - Talk at a conference/colloquium.
% - Talk length is about 20min.
% - Style is ornate.

\mode<presentation>
{
  \usetheme{Warsaw}
  % or ...

  %\setbeamercovered{transparent}
  % or whatever (possibly just delete it)
}

\usepackage[english]{babel}
\usepackage{listings}
\usepackage{ulem}
\usepackage{color}
\usepackage{alltt}
\usepackage{tikz}
\usepackage{pgfplots}

\usepackage[utf8x]{inputenc}

\newcommand\redsout[1]{{\color{red}\sout{\hbox{\color{black}{#1}}}}}

% or whatever

% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{The Efficient Handling of Guards in the Design of RPython's Tracing JIT}

\author[David Schneider, Carl Friedrich Bolz]{David Schneider \and \emph{Carl Friedrich Bolz}}
% - Give the names in the same order as the appear in the paper.
% - Use the \inst{?} command only if the authors have different
%   affiliation.

\institute[Heinrich-Heine-Universität Düsseldorf]{
Heinrich-Heine-Universität Düsseldorf, STUPS Group, Germany
}

\date{2012 VMIL, 21st of October, 2012}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
%\AtBeginSubsection[]
%{
%  \begin{frame}<beamer>
%    \frametitle{Outline}
%    \tableofcontents[currentsection,currentsubsection]
%  \end{frame}
%}

% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

%\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

%\section{Introduction}

\begin{frame}
  \frametitle{Tracing JITs Compile by Observing an Interpreter}
  \begin{itemize}
      \item VM contains both an interpreter and the tracing JIT compiler
      \item JIT works by observing and logging what the interpreter does
      \item for interesting, commonly executed code paths
      \item produces a linear list of operations (trace)
  \end{itemize}
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/loop01}
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/loop02}
\end{frame}

\begin{frame}
  \frametitle{Guards}
  \begin{itemize}
      \item Points of control flow divergence are marked with guards
      \item Operations that check whether conditions are still true
      \item When a guard fails, execution of the trace stops and continues in the interpreter
      \pause
      \item \emph{This talk:} technology and design decisions of guards
      \pause
      \begin{block}{Guard Characteristics}
          \begin{itemize}
              \item lots of them, up to 20\% guards
              \item most never fail
              \item need big information attached to them
          \end{itemize}
      \end{block}
  \end{itemize}
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/loop03}
\end{frame}

\begin{frame}
  \frametitle{Inlining}
  Tracing automatically does (potentially deep) inlining
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/loop04}
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/loop05}
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/loop06}
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/loop07}
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/loop08}
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/loop09}
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/loop10}
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/loop11}
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/loop12}
\end{frame}

% this talk wants to go over a lot of details that are usually glossed over as
% "easy" when tracing JITs are introduced.

\begin{frame}
  \frametitle{Bridges}
  \begin{itemize}
      \item When a trace fails often, it becomes worth to attach a new trace to it
          \item This is called a bridge
          \item The bridge is attached by patching the guard machine code
          \item when this guard fails in the future, the new trace is executed instead
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{RPython and PyPy}
  \begin{itemize}
      \item Context: RPython
      \item a generic tracing JIT, applicable to many languages
      \item main use: PyPy, an efficient Python interpreter
  \end{itemize}
\end{frame}

%\section{High-Level}

\begin{frame}
  \frametitle{Symbolic Frame Capturing}
  \begin{itemize}
      \item Guard can fail deep inside inlined function
      \item when going back to the interpreter, call stack needs to be re-created
      \item done with the help of symbolic frame stacks
      \item these show how trace variables fill the to-be-built stack frames
  \end{itemize}
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/loop07}
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/framechain1}
\end{frame}

\begin{frame}
  \frametitle{Symbolic Frame Compression}
  \begin{itemize}
      \item There are \emph{a lot of} guards
      \item Naively storing symbolic frames would be costly in terms of memory
      \item need to store them compactly
      \item observation: from one guard to the next, the non-top stack frames don't change
      \item share these between subsequent guards
      \pause
      \item also need a byte-saving binary representation, but that's just boring work
  \end{itemize}
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/loop07}
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/framechain1}
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/loop08}
\end{frame}

\begin{frame}
  \includegraphics[scale=0.4]{figures/framechain2}
\end{frame}

\begin{frame}
  \frametitle{Interaction with Optimization}
  \begin{itemize}
      \item Some optimizations make it necessary to store extra information in symbolic frames
      \pause
      \item examples:
          \begin{itemize}
              \item allocation removal (need to allocate objects before resuming)
              \item delayed heap stores (need to do stores before resuming interpreter)
          \end{itemize}
      \item can be compressed using similar techniques
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Emitting Guards}
  Guards are compiled as
  \begin{itemize}
    \item quick Check if the condition holds
    \item and a mapping of machine locations to JIT-variables % indirection using the fail-boxes
  \end{itemize}
  \pause
  In case of failure
  \begin{itemize}
    \item execution jumps to shared compensation code, decodes and stores mapping
    \item returns to interpreter that rebuilds state
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Compiling a Trace}
  \begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{figures/loop.pdf}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Compiling a Bridge}
  \begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{figures/bridge_compiled.pdf}
  \end{figure}
\end{frame}
\begin{frame}
  \frametitle{Patching Guards for Bridges}
  \begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{figures/bridge_patched.pdf}
  \end{figure}
\end{frame}
\begin{frame}[t,fragile]
\pgfplotsset{tick label style={font=\tiny\bfseries},
label style={font=\small},
legend style={font=\tiny}
}
    \frametitle{Guard Failure Rates}
    \begin{figure}
        \centering
        \begin{tikzpicture}
            \begin{axis}[
            xlabel= Guards by failures,
            ylabel=Relative \# of failures,
            xtick=\empty,
            ytick=\empty,
            ]
                \input{figures/data}
            \end{axis}
        \end{tikzpicture}
    \end{figure}
\end{frame}
\begin{frame}[t,fragile]
\pgfplotsset{tick label style={font=\tiny\bfseries},
label style={font=\small},
legend style={font=\tiny}
}
    \frametitle{Guard Failure Rates / Go Benchmark}
    \begin{figure}
        \centering
        \begin{tikzpicture}
            \begin{axis}[
            xlabel= Guards by failures,
            ylabel=Relative \# of failures,
            xtick=\empty,
            ytick=\empty,
            ]
                \input{figures/go_data}
            \end{axis}
        \end{tikzpicture}
    \end{figure}
\end{frame}

%\section{Evaluation}

%as in paper
%fancy graphs
%something about execution speed
% <demo, if there is time>
\end{document}