### 2014-06-05 talk/pycon2010/crossinterp/talk.tex
Attempt to simplify the Memory Requirements sections into just some text in Future Work.

\documentclass[utf8x, 14pt]{beamer}

\mode<presentation>
{
  \usetheme{Warsaw}
}

\usepackage[english]{babel}

\usepackage[utf8x]{inputenc}
\usepackage{tikz}

\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{color}

\title{How to write cross-interpreter programs}

\author{Maciej Fijałkowski}

\institute[merlinux GmbH]
{ merlinux GmbH }

\date{Pycon 2010, February 19th 2010, Atlanta}

\begin{document}

\begin{frame}
  \titlepage
  \begin{figure}
    \begin{tabular}{c c c}
    \includegraphics[width=.20\textwidth]{../common/pypy-logo.png}
    &
    \hspace{2cm}
    &
    \includegraphics[width=.18\textwidth]{../common/merlinux-logo.png}
    \end{tabular}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{This talk}
  \begin{itemize}
    \item most people only target CPython (or Jython or IronPython)
    \item sometimes, you want your program to run on each of those
    \item libraries are more often cross-interpreter
      \pause
    \item hope to give you more reasons tomorrow
      \pause
    \item won't talk about py3k
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{My background}
  \begin{itemize}
    \item worked a lot with PyPy compatibility issues
    \item helped to port twisted, django and other projects to run on PyPy
    \item a lot of cooperation with Jython people
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Have a way to verify compatibility}
  \begin{itemize}
    \item extensive test suite
    \item good coverage
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Exceptions}
  \begin{itemize}
    \item TypeError vs AttributeError change often between implementations,
      even CPython versions
    \item don't rely on exception string messages (they may differ)
  \end{itemize}
  \begin{verbatim}
try:
  ...
except ImportError, ie:
  if str(ie) != '...':
    raise
  \end{verbatim}
  \pause
  \begin{itemize}
    \item also means - don't use doctests
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Subclasses of builtin types}
  \begin{itemize}
    \item in general overridden methods on subclassed builtin types are not 
      invoked by preexisting other methods
      \begin{verbatim}
class d(dict):
  def __getitem__(self, e):
    ...
      \end{verbatim}
    \item would {\ttfamily keys()} go via this getitem?
    \item tests are your friend
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Access to 3rdy party libraries}
  \begin{itemize}
    \item there is no good story here
    \item ctypes based access is going to be supported by all Pythons
    \item are there pure Python replacements/options?
    \item separate out dependencies/especially optional ones
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Don't rely on refcounting}
  \begin{itemize}
    \item example
      \begin{verbatim}
open('x', 'w').write('stuff')
      \end{verbatim}
    \item on refcounting, flushes file immediately
    \item on any other GC, it might be deferred for
      a while
    \item the single most-common problem when porting twisted to PyPy
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{\_\_del\_\_}
  \begin{itemize}
    \item resurrection on CPython will call {\ttfamily \_\_del\_\_} multiple times,
      other Pythons exactly once
    \item cycles with {\ttfamily \_\_del\_\_}s are not collected by CPython,
      PyPy breaks them randomly instead
    \item in PyPy and Jython {\ttfamily \_\_del\_\_} cannot be attached to classes after creation
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Use new-style classes}
  \begin{itemize}
    \item 3.x ready
    \item much faster on PyPy, too
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{{\ttfamily sys.prefix}}
  XXX think about
  \begin{itemize}
    \item implementations may have different installation layouts
    \item open issue, at least for PyPy, has compatibility consequences with
  setuptools
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{IO bytes vs unicode}
  \begin{itemize}
    \item convert/decode as soon as possible, keep text and bytes apart
    \item for 2.x Pythons use str for bytes and unicode for text
    \item the distinction is deeper in 3.x ({\ttfamily str} is {\ttfamily unicode}, {\ttfamily bytes} exist with
  slightly different interface than old {\ttfamily str})
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Don't concatenate strings}
  \begin{itemize}
    \item use {\ttfamily "".join(...)}
    \item if you care about performance, try this and {\ttfamily cStringIO}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Obscure corners}
  \begin{itemize}
    \item non-string keys in type dictionaries
    \item introspection results, implementation objects (e.g. builtin methods
      etc), may have different types
    \item exact naming of things (like list-comprehension variable)
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Questions?}
  \begin{itemize}
    \item http://morepypy.blogspot.com
    \item http://pypy.org
  \end{itemize}
\end{frame}

\end{document}