### 2014-10-01 talk/badhonnef2007/badhonnef.tex
More slides

\documentclass{llncs}

\usepackage{makeidx}
\usepackage{graphicx}

\begin{document}

\pagestyle{headings}

\title{A Flexible Prolog Interpreter in Python}

%\titlerunning{XXX}     % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used

\author{Carl Friedrich Bolz}

\authorrunning{Bolz}   % abbreviated author list (for running head)

%%%% modified list of authors for the TOC (add the affiliations)
\tocauthor{Bolz (D\"usseldorf)}

\institute{ Lehrstuhl Softwaretechnik und Programmiersprachen\\
Institut f\"{u}r Informatik, Universit\"at D\"usseldorf, Germany\\
  \email{ cfbolz@gmx.de}}

\maketitle

\begin{abstract}

Efficient virtual machines for the Prolog programming language (and also for
other dynamic highlevel languages) are usually large and intricate pieces of
software implemented in a low-level language such as C that are extensively
optimized over many years and are therefore difficult to change in many
respects. Due to their inherent complexity it is hard to develop them further,
to maintain them or to use them to experiment with language extensions or new
built-ins.

Early implementation decisions, such as choice of garbage collector, are
extremely hard to change afterwards. In addition, it is also much harder to
perform advanced optimizations or transformations (such as partial evaluation)
on the virtual machine itself due to the low level of the implementation
language. Furthermore it is (by definition) impossible to port them to a non-C
platform such as the Java Virtual Machine or the .NET framework Therefore the
numerous implementations of a Prolog virtual machine integrated into one of
these environments are all new implementations which have to be maintained
themselves and therefore all have different levels of maturity and slightly
different sets of built-ins available.

Using a higher-level language than C addresses many of these problems. It
allows the language implementor to work at a higher level of abstraction, which
eases implementation, maintenance and extensions of the virtual machine and
reduces the size of the code base.  Furthermore, advanced optimizations and
transformations can be performed. To regain usable performance the
implementation needs then to be translated to a low-level language again.

The goal of the PyPy project is to support the implementation of interpreters
in the high-level language Python. Python is an object oriented dynamic
language which is becoming more and more popular for scripting use as well as
for “real” applications. PyPy provides a translation toolchain that helps to
translate interpreters written in Python to a low-level language such as C and
the .NET Intermediate Language. To make this translation work, these
interpreters have to be written in RPython, which is a proper subset of Python
chosen in such a way that it is amenable to analysis.

In this talk we give an quick overview of the PyPy project and present a Prolog
interpreter written in RPython, exploring the possibilities of implementing a
well-known declarative language in a high-level language (as opposed to C). One
of the main goals of the implementation was to keep it as simple and as
extensible as possible (even at the cost of performance). The PyPy tool suite
is used to reach a reasonable level of performance and to ensure portability to
various platforms.\footnote{This research was partially supported by the EU funded
 project: IST 004779 PyPy (PyPy: Implementing Python in Python).} \\

\end{abstract}

% ---- Bibliography ----
%\begin{small}
\bibliographystyle{abbrv}
\bibliography{badhonnef}
%\end{small}

\end{document}